---
title: "Second-anchor overlap with Xiang epiblast enhancers (GRCm38), by cluster group"
output:
  html_document:
    df_print: paged
---

For each interaction we take the **second** genomic interval encoded in `interaction_ID` (format `chr_start_end...chr_start_end`) and ask how many of these regions overlap the epiblast enhancer calls from Xiang et al. (GRCm38 BED). Only `interaction_ID`, `EM_cluster_group`, and `Louvain_cluster_group` are read from `interaction_pairs_uncapped.txt`. Aside from subsetting by the cluster group for each panel, **no other filtering** is applied.

**What counts as overlap:** each second-anchor interval is compared to the Xiang BED using `GenomicRanges::countOverlaps` (default **any** overlap between intervals on the same chromosome, strand ignored). A region is counted in the **Overlap** slice if it shares **at least one base pair** with **at least one** enhancer interval. This is a binary yes/no per interaction; it is **not** the fraction of the query interval covered by enhancers, nor a minimum overlap length beyond that single-base requirement.


Loading packages and options

```{r setup, message=FALSE}
library(tidyverse)
library(GenomicRanges)
library(patchwork)
options(scipen = 999)
theme_set(theme_bw(base_size = 16))
dir.create("../output/plots", showWarnings = FALSE, recursive = TRUE)
```


Cluster groups to compare (same labels for EM and Louvain)

```{r cluster_groups}
cluster_groups <- c(
  "active P - active E",
  "inactive P - primed E",
  "inactive P - unmarked PIR"
)
```


Read only the identifier and cluster columns; load Xiang enhancers once

```{r import}
read_tsv(
  "../data/interaction_pairs_uncapped.txt",
  col_select = all_of(c("interaction_ID", "EM_cluster_group", "Louvain_cluster_group")),
  show_col_types = FALSE
) -> pairs_min

read.table(
  "../data/Xiang_epiblast_enhancers_GRCm38.bed",
  col.names = c("chr", "start", "end"),
  stringsAsFactors = FALSE
) -> xiang_bed

gr_xiang <- GRanges(
  seqnames = xiang_bed$chr,
  ranges = IRanges(start = xiang_bed$start, end = xiang_bed$end)
)
```


Parse each second anchor, drop rows that do not match the expected pattern, then test overlap (one logical per successfully parsed interval)

```{r overlap_fn}
overlap_with_xiang <- function(interaction_ids) {
  parsed <- tibble(interaction_ID = interaction_ids) %>%
    mutate(second = str_split(interaction_ID, fixed("..."), simplify = TRUE)[, 2]) %>%
    tidyr::extract(
      second,
      into = c("chr", "start", "end"),
      regex = "^([^_]+)_([^_]+)_([^_]+)$",
      convert = TRUE
    ) %>%
    drop_na(chr, start, end)
  if (nrow(parsed) == 0L) {
    return(logical(0))
  }
  gr_q <- GRanges(
    seqnames = paste0("chr", parsed$chr),
    ranges = IRanges(start = parsed$start, end = parsed$end)
  )
  # any overlap on the same chromosome (>= 1 shared bp); not %-coverage of the query
  countOverlaps(gr_q, gr_xiang, ignore.strand = TRUE) > 0
}

make_overlap_pie <- function(overlap_hits, plot_title) {
  n_tot <- length(overlap_hits)
  if (n_tot == 0L) {
    return(
      ggplot() +
        theme_void(base_size = 16) +
        labs(
          title = plot_title,
          caption = "Total regions: 0"
        ) +
        theme(
          plot.title = element_text(hjust = 0.5, size = 11),
          plot.caption = element_text(hjust = 0.5, size = 10, margin = margin(t = 8))
        ) +
        annotate("text", x = 0, y = 0, label = "No interactions", size = 4)
    )
  }
  n_ov <- sum(overlap_hits)
  n_no <- sum(!overlap_hits)
  pie_tbl <- tibble(
    status = factor(c("Overlap", "No overlap"), levels = c("Overlap", "No overlap")),
    n = c(n_ov, n_no)
  ) %>%
    mutate(pct = 100 * n / sum(n))

  ggplot(pie_tbl, aes(x = "", y = n, fill = status)) +
    geom_bar(
      stat = "identity",
      width = 1,
      color = "white",
      linewidth = 0.2,
      position = position_stack(reverse = TRUE)
    ) +
    coord_polar(theta = "y", start = 0, direction = 1) +
    scale_fill_manual(
      values = c("No overlap" = "grey75", "Overlap" = "grey25"),
      name = NULL
    ) +
    geom_text(
      aes(label = paste0(round(pct, 1), "%")),
      position = position_stack(reverse = TRUE, vjust = 0.5),
      size = 3.2
    ) +
    labs(
      title = plot_title,
      caption = paste0("Total regions: ", format(n_tot, big.mark = ","))
    ) +
    theme_void(base_size = 16) +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, size = 11, face = "plain"),
      plot.caption = element_text(hjust = 0.5, size = 10, margin = margin(t = 8))
    )
}
```


### EM clustering

```{r pies_em, fig.width = 12, fig.height = 4.6}
plots_em <- map(cluster_groups, function(g) {
  ids <- pairs_min %>%
    filter(EM_cluster_group == g) %>%
    pull(interaction_ID)
  hits <- overlap_with_xiang(ids)
  make_overlap_pie(hits, paste0("EM cluster group ", g))
})
p_em <- wrap_plots(plots_em, ncol = 3)
p_em

dir.create("../output/plots", showWarnings = FALSE, recursive = TRUE)
ggsave(
  "../output/plots/SuppFigure4b_Xiang_enhancer_overlap_pies_EM.svg",
  p_em,
  width = 12,
  height = 4.6,
  units = "in",
  device = "svg"
)

ggsave(
  "../output/plots/SuppFigure4b_Xiang_enhancer_overlap_pies_EM.png",
  p_em,
  width = 12,
  height = 4.6,
  units = "in",
  device = "png"
)
```


### Louvain clustering

```{r pies_louvain, fig.width = 12, fig.height = 4.6}
plots_lou <- map(cluster_groups, function(g) {
  ids <- pairs_min %>%
    filter(Louvain_cluster_group == g) %>%
    pull(interaction_ID)
  hits <- overlap_with_xiang(ids)
  make_overlap_pie(hits, paste0("Louvain cluster group ", g))
})
p_lou <- wrap_plots(plots_lou, ncol = 3)
p_lou

ggsave(
  "../output/plots/SuppFigure4b_Xiang_enhancer_overlap_pies_Louvain.svg",
  p_lou,
  width = 12,
  height = 4.6,
  units = "in",
  device = "svg"
)

ggsave(
  "../output/plots/SuppFigure4b_Xiang_enhancer_overlap_pies_Louvain.png",
  p_lou,
  width = 12,
  height = 4.6,
  units = "in",
  device = "png"
)
```
