---
title: "Alternative clustering for Plex data"
output: html_notebook
---

We have been asked by a reviewer to provide details on why we did what we did in terms of clustering. Here, I want to try out a different clustering approach to see if we get overlapping clustering and how sensitive the cluster borders are to different paramters.

I thought I could try and use the Seurat workflow and 'pretend' that the magic table is sc RNAseq data.
I have mainly been following this tutorial https://satijalab.org/seurat/articles/pbmc3k_tutorial.html

Loading libraries

```{r}
library(Seurat)
library(tidyverse)
```

Importing the interaction pair data

```{r}
read_tsv("../data/umap_mark_data_uncapped.txt") -> marks_import
head(marks_import)
colnames(marks_import)
```
Retaining only the columns that should go into clustering

```{r}
marks_import %>% 
  select(-contains("UMAP"), -P_ID, -PIR_ID, -P_gene, -PIR_gene, -contains("Input"), -cluster_ID, -cluster_group) %>% # the labels and marks I don't want
  select(-contains("DNMT"), -contains("ES"), -contains("KO")) -> marks # limiting to Epi and TET and WT
head(marks)
```
Transposing this so it can be made into a Seurat object

```{r}
marks %>% 
  column_to_rownames(var = "interaction_ID") %>% 
  t() %>% 
  as.data.frame() -> marks_transposed

head(marks_transposed)
```


Making this into a Seurat object

```{r}
CreateSeuratObject(counts = marks_transposed, assay = "mark") -> marks_Seurat
```

LogNormalising the values

```{r}
marks_normalised <- NormalizeData(marks_Seurat)
```

Scaling the data

```{r}
all.marks <- rownames(marks_normalised)
marks_scaled <- ScaleData(marks_normalised, features = all.marks)
```

performing linear dimensional reduction

```{r}
marks_PCA <- RunPCA(marks_scaled, features = all.marks)
```

The warnings here are ok, it just complains that it doesn't have enough dimensions and automatically reduces them to 17.
```{r}
DimPlot(marks_PCA, reduction = "pca", raster = FALSE)
```

```{r}
#DimHeatmap(marks_PCA, dims = 1:5, balanced = TRUE)
```
```{r}
ElbowPlot(marks_PCA)
```
It's hard to determine a cutoff for PCAs to use, I tried 8, 9 and 12 to check the clustering.

Clustering

-> KNN graph based on euclidean distance in PCA space with edge weights between samples by Jaccard similarity (overlap in local neighbourhoods)
-> Louvain clustering 

```{r}
marks_KNN <- FindNeighbors(marks_PCA, dims = 1:12)
marks_Louvain <- FindClusters(marks_KNN, resolution = 1)
```

I ran several iterations with different parameters to check what made the most sense with the UMAP representation.
dim 9, resolution 0.8
dim 9, resolution 1.2
dim 15, resolution 1
dim 12, resolution 1 (this is the one I settled on)


Look at cluster IDs

```{r}
Idents(marks_Louvain) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "interaction_ID") %>% 
  rename(Louvain_cluster_ID = ".") -> Louvain_clusters
head(Louvain_clusters)
```

Now incorporating this info in the original table

```{r}
marks_import %>% 
  left_join(Louvain_clusters) -> marks_2
head(marks_2)
```

Plotting the original clusters

```{r}
marks_2 %>% 
  slice_sample(prop = 0.1) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = cluster_ID)) +
  geom_point(size = 0.5) +
  theme_bw(base_size = 14) +
  ggtitle("EM clusters") +
  coord_cartesian(xlim = c(-10, 12), ylim = c(-10, 8)) +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))
```
Plotting the Louvain clusters

```{r}
marks_2 %>% 
  slice_sample(prop = 0.05) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = Louvain_cluster_ID)) +
  geom_point(size = 0.5) +
  geom_point(data = marks_2 %>% slice_sample(prop = 0.05) %>% filter(Louvain_cluster_ID == 9), size = 0.5, colour = "black") +
  theme_bw(base_size = 14) +
  ggtitle("Louvain clusters") +
  coord_cartesian(xlim = c(-10, 12), ylim = c(-10, 8)) +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))
```

Okay, let's explore these a bit more.

How many regions per cluster?

```{r}
marks_2 %>% 
  group_by(Louvain_cluster_ID) %>% 
  count()
```
Looking at marks per cluster

```{r}
marks_2 %>% 
  select(-UMAP1, -UMAP2, -P_ID, -PIR_ID, -P_gene, -PIR_gene, -contains("DNMT"), -contains("ES"), -contains("Input"), -contains("KO"), -contains("CHiC")) %>% 
  select(interaction_ID, cluster_ID, cluster_group, Louvain_cluster_ID, everything()) %>% 
  pivot_longer(5:last_col(), names_to = "condition", values_to = "values") %>% 
  separate(condition, into = c("region", "mark")) -> marks_TET_Epi_long
```
The warning here is fine, I deliberately discard the additional pieces.

Clusters to exclude (the ones with fewer than 1000 data points)

```{r}
excluded_clusters <- c("21", "22", "23", "24", "25", "26", "27", "28", "29", "30")
```


Plotting boxplots 

```{r, fig.height=6, fig.width = 4}
# marks_TET_Epi_long %>% 
#   filter(mark != "mC") %>% 
#   filter(!Louvain_cluster_ID %in% excluded_clusters) %>% 
#   ggplot(aes(Louvain_cluster_ID, values, fill = region)) +
#   geom_boxplot(outlier.shape = NA) +
#   ylim(0, 8) +
#   theme_bw(base_size = 14) +
#   facet_grid(rows = "mark")
```


Boxplots per cluster

```{r, fig.height=17, fig.width = 10}
marks_TET_Epi_long %>% 
  filter(mark != "mC") %>% 
  filter(!Louvain_cluster_ID %in% excluded_clusters) %>% 
  ggplot(aes(mark, values, fill = region)) +
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 8) +
  theme_bw(base_size = 14) +
  facet_grid(rows = "Louvain_cluster_ID")
```
I then summarised the clusters according to their marks (and their presumed biological function).


Summarised clusters

```{r}
cluster_group_names <- c("active P - unmarked PIR",
                        "inactive P - low primed E",
                        "PC P - unmarked PIR",
                        "inactive P - unmarked PIR",
                        "inactive P - low primed E, CTCF",
                        "primed E",
                        "poised E",
                        "active P - active P",
                        "active P - active E"
                        )
cluster_group_colours <- c("#b785bb", "#53c2b3", "#bf6c6c", "#b2b5b5", "#741113", "#eaca48", "#ca3526", "#91c740", "#252d37")
names(cluster_group_colours) <- cluster_group_names
```

Plotting original summarised marks

```{r}
marks_2 %>% 
  slice_sample(prop = 0.1) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = cluster_group)) +
  geom_point(size = 0.5) +
  theme_bw(base_size = 14) +
  ggtitle("EM clusters") +
  coord_cartesian(xlim = c(-10, 12), ylim = c(-10, 8)) +
  scale_colour_manual(values = cluster_group_colours) +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))
```


Summarising Louvain clusters
(This one is for dim 12, resolution 1)

```{r}
P_unmarkedPIR <- c("2", "5", "13", "15")
inactive_lowprimed <- c("9")
PC_P_unmarked <- c("0", "12", "17")
inactive_unmarked <- c("3", "4", "7", "9", "10", "14", "16")
inactive_lowprimedCTCF <- c("1", "19")
primed <- c("20", "6")
poised <- c( "18")
p_p <- c("11")
active_activeE <- c("8")
```


```{r}
marks_2 %>% 
  filter(!Louvain_cluster_ID %in% excluded_clusters) %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_ID %in% P_unmarkedPIR ~ "active P - unmarked PIR",
                                           Louvain_cluster_ID == inactive_lowprimed ~ "inactive P - low primed E",
                                           Louvain_cluster_ID %in% PC_P_unmarked ~ "PC P - unmarked PIR",
                                           Louvain_cluster_ID %in% inactive_unmarked ~ "inactive P - unmarked PIR",
                                           Louvain_cluster_ID %in% inactive_lowprimedCTCF ~ "inactive P - low primed E, CTCF",
                                           Louvain_cluster_ID == primed ~ "primed E",
                                           Louvain_cluster_ID %in% poised ~ "poised E",
                                           Louvain_cluster_ID == p_p ~ "active P - active P",
                                           Louvain_cluster_ID %in% active_activeE ~ "active P - active E")) -> marks_3
```

Plotting Louvain summarised marks

```{r}
marks_3 %>% 
  slice_sample(prop = 0.05) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = Louvain_cluster_group)) +
  geom_point(size = 0.5) +
  theme_bw(base_size = 14) +
  ggtitle("Louvain clusters") +
  coord_cartesian(xlim = c(-10, 12), ylim = c(-10, 8)) +
  scale_colour_manual(values = cluster_group_colours) +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5))
```

Some clusters are strong and essentially pop up in all cluster conditions:

- polycomb promoters
- poised enhancers
- inactive P - CTCF
- active P - active P, although this one mixes easily with the active P active E cluster despite being quite defined on the UMAP
- we then have the various clusters that form the 'body' of the UMAP which aren't clearly defined. They have various levels of K27ac at P and PIR and are therefore labelled as low primed enhancers, active enhancers, active and inactive promoters. Maybe we should be clearer about this in the text.

