---
title: "Chromatin marks plots"
output: html_notebook
---


Looking at chromatin marks for each of the clusters 

Defining source, libraries and options

```{r}
source("cluster_analysis_functions.R")
library(easystats)
library(broom)
options(scipen = 0)
```


Importing datasets

The marks table was created for capped and uncapped values (the latter only to be used for colouring the umaps).
Linkings of marks to clusters happens in the script series umap_20_01_20_unccapped_marks and was exported from there.

```{r}
#read_tsv("../data/clusters_20_capped_marks_220517.txt") -> clusters_capped_marks
read_tsv("../data/clusters_20_uncapped_marks_220517.txt") -> clusters_uncapped_marks
#read_tsv("../data/pluripotency_exit_RNA_Epi_and_EB_log2RPKM.txt") -> expression_import
```

```{r}
clusters_uncapped_marks %>% 
  rename_with(~gsub("H3K27Ac", "H3K27ac",.)) -> marks
colnames(marks)
```

Getting some stats for Figure 3

Figure 3a, Dnmt TKO - WT

```{r}
marks %>% 
  filter(cluster_ID == "cluster16") %>% 
  select(contains("Epi")) -> cluster16_Epi

head(cluster16_Epi)
```
The data doesn't look normal. Let's check with Shapiro-Wilk test

```{r}
ks.test(cluster16_Epi$oe_H3K27ac_DNMT_WT_Epi, 'pnorm')
```
Because normality isn't met, use a non-parametric test although at that sample size it doesn't matter -> Mann Whitney U test

K27ac_DNMT

```{r}
wilcox.test(cluster16_Epi$oe_H3K27ac_DNMT_WT_Epi, cluster16_Epi$oe_H3K27ac_DNMT_KO_Epi) %>% 
  tidy() %>% 
  mutate(comparison = "K27ac_DNMT WT vs KO") %>% 
  relocate(comparison, .before = statistic) -> K27ac_DNMT_stats

K27ac_DNMT_stats
```

Effect size

```{r}
cohens_d(cluster16_Epi$oe_H3K27ac_DNMT_WT_Epi, cluster16_Epi$oe_H3K27ac_DNMT_KO_Epi) 
interpret_cohens_d(d = 0.12, rules = "cohen1988")
```


K4me1_DNMT

```{r}
wilcox.test(cluster16_Epi$oe_H3K4me1_DNMT_WT_Epi, cluster16_Epi$oe_H3K4me1_DNMT_KO_Epi) %>% 
  tidy() %>% 
  mutate(comparison = "K4me1_DNMT WT vs KO") %>% 
  relocate(comparison, .before = statistic) -> K4me1_DNMT_stats

K4me1_DNMT_stats
```
Effect size

```{r}
cohens_d(cluster16_Epi$oe_H3K4me1_DNMT_WT_Epi, cluster16_Epi$oe_H3K4me1_DNMT_KO_Epi) 
interpret_cohens_d(d = 0.75, rules = "cohen1988")
```

Figure 3a: TET WT - KO

K27ac_TET effect size

```{r}
cohens_d(cluster16_Epi$oe_H3K27ac_TET_WT_Epi, cluster16_Epi$oe_H3K27ac_TET_KO_Epi) 
interpret_cohens_d(d = 0.58, rules = "cohen1988")
```
K4me1_TET effect size

```{r}
cohens_d(cluster16_Epi$oe_H3K4me1_TET_WT_Epi, cluster16_Epi$oe_H3K4me1_TET_KO_Epi) 
interpret_cohens_d(d = 0.48, rules = "cohen1988")
```

------------------------------------------------------------------------------

## Ok, needing a more generalised solution:

### Functions

Here is a function that calculates Cohen's d between WT and KO

```{r}
calculate_cohens_d <- function(data, variable_prefix, separator = "_") {
  wt_columns <- grep(paste0("^", variable_prefix, ".*", "WT"), names(data), value = TRUE)
  ko_columns <- grep(paste0("^", variable_prefix, ".*", "KO"), names(data), value = TRUE)

  if (length(wt_columns) != length(ko_columns)) {
    stop("Unequal number of WT and KO columns")
  }

  result <- tibble()
  for (i in seq_along(wt_columns)) {
    wt_col <- data[[wt_columns[i]]]
    ko_col <- data[[ko_columns[i]]]
    comparison_name <- wt_columns[i]
    comparison_name <- gsub("_WT_Epi", "", comparison_name)
    cohens_d_value <- cohens_d(wt_col, ko_col)
    result <- bind_rows(result, tibble(Comparison = comparison_name, Cohens_d = cohens_d_value))
  }

  return(result)
}
```

Here is a function that runs the calculate Cohen's d function across all the histone marks I need

```{r}

calculate_cohens_d_multiple <- function(prefixes_to_calculate) {
  
  result_nested <- calculate_cohens_d(input_tibble, variable_prefix = prefixes_to_calculate)
  result_nested %>% 
    unnest(cols = c(Cohens_d)) -> results
  return(results)
  
}
```



### Cohen's d for cluster16_ Epi

```{r}
prefixes_to_calculate <- c("bait_H3K27ac", "oe_H3K27ac", "bait_H3K4me1", "oe_H3K4me1", "bait_H3K27me3", "oe_H3K27me3")
input_tibble <- cluster16_Epi

map(prefixes_to_calculate, calculate_cohens_d_multiple) %>% 
  bind_rows() %>% 
  separate(Comparison, into = c("fragment", "mark", "genotype")) -> cluster16_cohensd

cluster16_cohensd
```


### Cohen's d for cluster12_ Epi

Subsetting the cluster

```{r}
marks %>% 
  filter(cluster_ID == "cluster12") %>% 
  select(contains("Epi")) -> cluster12_Epi

head(cluster12_Epi)
```
Calculating Cohen's d

```{r}
prefixes_to_calculate <- c("bait_H3K27ac", "oe_H3K27ac", "bait_H3K4me1", "oe_H3K4me1", "bait_H3K27me3", "oe_H3K27me3")
input_tibble <- cluster12_Epi

map(prefixes_to_calculate, calculate_cohens_d_multiple) %>% 
  bind_rows() %>% 
  separate(Comparison, into = c("fragment", "mark", "genotype")) -> cluster12_cohensd

cluster12_cohensd
```

### Cohen's d for cluster2_ Epi

Subsetting the cluster

```{r}
marks %>% 
  filter(cluster_ID == "cluster2") %>% 
  select(contains("Epi")) -> cluster2_Epi

head(cluster2_Epi)
```

Calculating Cohen's d

```{r}
prefixes_to_calculate <- c("bait_H3K27ac", "oe_H3K27ac", "bait_H3K4me1", "oe_H3K4me1", "bait_H3K27me3", "oe_H3K27me3")
input_tibble <- cluster2_Epi


map(prefixes_to_calculate, calculate_cohens_d_multiple) %>% 
  bind_rows() %>% 
  separate(Comparison, into = c("fragment", "mark", "genotype")) -> cluster2_cohensd

cluster2_cohensd
```
### Cohen's d for cluster17_ Epi

Subsetting the cluster

```{r}
marks %>% 
  filter(cluster_ID == "cluster17") %>% 
  select(contains("Epi")) -> cluster17_Epi

head(cluster17_Epi)
```

Calculating Cohen's d

```{r}
prefixes_to_calculate <- c("bait_H3K27ac", "oe_H3K27ac", "bait_H3K4me1", "oe_H3K4me1", "bait_H3K27me3", "oe_H3K27me3")
input_tibble <- cluster17_Epi


map(prefixes_to_calculate, calculate_cohens_d_multiple) %>% 
  bind_rows() %>% 
  separate(Comparison, into = c("fragment", "mark", "genotype")) -> cluster17_cohensd

cluster17_cohensd
```

## Visualisation

Preparing the data frame for plotting

```{r}
cluster16_cohensd %>% 
  rename(`Active E (16)` = Cohens_d) %>% 
  select(mark, fragment, genotype, `Active E (16)`) -> cluster16_temp

cluster12_cohensd %>% 
  rename(`Primed E (12)` = Cohens_d) %>% 
  select(mark, fragment, genotype, `Primed E (12)`) -> cluster12_temp

cluster2_cohensd %>% 
  rename(`Polycomb (2)` = Cohens_d) %>% 
  select(mark, fragment, genotype, `Polycomb (2)`) -> cluster2_temp

cluster17_cohensd %>% 
  rename(`Poised E (17)` = Cohens_d) %>% 
  select(mark, fragment, genotype, `Poised E (17)`) -> cluster17_temp

cluster12_temp %>% 
  left_join(cluster16_temp) %>% 
  left_join(cluster17_temp) %>% 
  left_join(cluster2_temp) %>% 
  mutate(fragment = gsub("bait", "P", fragment)) %>% 
  mutate(fragment = gsub("oe", "PIR", fragment)) %>% 
  mutate(mark = factor(mark, levels = c("H3K27me3", "H3K27ac", "H3K4me1")))  -> summary_cohensd_wide
  
summary_cohensd_wide %>% 
  pivot_longer(4:last_col(), names_to = "cluster", values_to = "Cohens_d") %>% 
  mutate(cluster = factor(cluster, levels = c("Active E (16)", "Primed E (12)", "Poised E (17)", "Polycomb (2)"))) %>% 
  mutate(Cohens_d = round(Cohens_d, digits = 2))-> summary_cohensd

summary_cohensd
```

Generating the heatmap

```{r, fig.width=7, fig.height=5}
summary_cohensd %>% 
  ggplot(aes(cluster, mark, fill = -Cohens_d)) +
  geom_tile() +
  geom_text(label = -(summary_cohensd$Cohens_d)) +
  facet_grid(rows = vars(genotype), cols = vars(fragment)) +
  #scale_fill_viridis_c(option = "turbo") + 
  scale_fill_distiller(palette = "RdYlBu", direction = -1, limits = c(-0.8, 0.8), na.value = "#d73026") +
  theme(axis.title = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5), legend.title = element_blank()) -> cohens_heatmaps

cohens_heatmaps
```
export

```{r}
ggsave("../output/plots/cohens_heatmaps.png", cohens_heatmaps, width = 7, height = 5)
ggsave("../output/plots/cohens_heatmaps.svg", cohens_heatmaps, width = 7, height = 5)
```


