---
title: "ES to Epi transition"
output: html_notebook
---

Looking at chromatin marks for each of the cluster group in the transition from ES to Epi.
This is an adapted script from June 22 but using the new cluster groups (EM 15 dim and Louvain)

Defining source

```{r}
source("cluster_analysis_functions.R")
```


Importing datasets

The marks table was created for capped and uncapped values (the latter only to be used for colouring the umaps).
Linkings of marks to clusters happens in the script grouping_clusters.Rmd and the 'magic tables' were exported from there.

```{r}
read_tsv("../data/interaction_pairs_capped.txt") -> clusters_capped_marks
read_tsv("../data/interaction_pairs_uncapped.txt") -> clusters_uncapped_marks
```


Tidying expression 

```{r}
expression_import %>% 
  distinct(Probe, .keep_all = TRUE) %>% 
  select(Probe,13:last_col()) %>% 
  rename_with(.cols = 2:last_col(), ~paste0("bait_RNA_",.)) %>% 
  rename_with(.cols = 2:last_col(), ~gsub("ES", "ES_0",.)) -> expression

head(expression)
```


Merging with marks data

```{r}
clusters_capped_marks %>% 
  left_join(expression, by = c("bait_gene" = "Probe")) %>% 
  relocate(cluster_ID, .after = interaction_ID) %>% 
  rename_with( ~gsub("H3K27Ac", "H3K27ac",.)) %>%  
  mutate(cluster_ID = factor(cluster_ID, levels = c("cluster0", "cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "cluster6", "cluster7", "cluster8", "cluster9", "cluster10", "cluster11", "cluster12", "cluster13", "cluster14", "cluster15", "cluster16", "cluster17", "cluster18", "cluster19"))) -> clusters_capped

clusters_uncapped_marks %>% 
  left_join(expression, by = c("bait_gene" = "Probe")) %>% 
  relocate(cluster_ID, .after = interaction_ID) %>% 
  rename_with( ~gsub("H3K27Ac", "H3K27ac",.)) %>%  
  mutate(cluster_ID = factor(cluster_ID, levels = c("cluster0", "cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "cluster6", "cluster7", "cluster8", "cluster9", "cluster10", "cluster11", "cluster12", "cluster13", "cluster14", "cluster15", "cluster16", "cluster17", "cluster18", "cluster19"))) -> clusters_uncapped

head(clusters_uncapped)
```

Making the comparison for TET WT ES vs EPI

```{r}
clusters_uncapped %>% 
  select(-contains("DNMT"), -contains("KO"), -contains("RNA")) %>% 
  wide_cluster_table_to_long() %>% 
  select(-line, -genotype, -day) -> TET_WT

head(TET_WT)
```

I'll have to adjust the y axis individually, that's why I'll be plotting these manually and then export. 

```{r, fig.width=20, fig.height=5}
TET_WT %>% 
  mutate(cluster_ID = gsub("cluster", "", cluster_ID)) %>% 
  mutate(cluster_ID = factor(cluster_ID, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19"))) %>% 
  filter(mark == "mC") %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ cluster_ID) +
  coord_cartesian(ylim = c(0, 100)) +
  ggtitle("TET WT mC") +
  theme(legend.position = "none", axis.title.x = element_blank(), plot.title = element_text(size = 14, hjust = 0.5))
```

Exporting

```{r}
ggsave("../output/plots/ES_to_Epi/TET_ES_to_Epi_mC.png", width = 20, height = 5, units = "in")
ggsave("../output/plots/ES_to_Epi/TET_ES_to_Epi_mC.svg", width = 20, height = 5, units = "in")
```

Only one row for CHiC

```{r, fig.width=20, fig.height=3}
TET_WT %>% 
  mutate(cluster_ID = gsub("cluster", "", cluster_ID)) %>% 
  mutate(cluster_ID = factor(cluster_ID, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19"))) %>% 
  filter(mark == "CHiC") %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5) +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ cluster_ID) +
  coord_cartesian(ylim = c(0,10)) +
  ggtitle("TET WT CHiC") +
  theme(legend.position = "none", axis.title.x = element_blank(), plot.title = element_text(size = 14, hjust = 0.5))
```
Exporting

```{r}
ggsave("../output/plots/ES_to_Epi/TET_ES_to_Epi_CHiC.png", width = 20, height = 3, units = "in")
ggsave("../output/plots/ES_to_Epi/TET_ES_to_Epi_CHiC.svg", width = 20, height = 3, units = "in")
```


Making the comparison for DNMT WT ES vs EPI

```{r}
clusters_uncapped %>% 
  select(-contains("TET"), -contains("KO"), -contains("RNA")) %>% 
  wide_cluster_table_to_long() %>% 
  select(-line, -genotype, -day) -> DNMT_WT

head(DNMT_WT)
```


I'll have to adjust the y axis individually, that's why I'll be plotting these manually and then export. 

```{r, fig.width=20, fig.height=5}
DNMT_WT %>% 
  mutate(cluster_ID = gsub("cluster", "", cluster_ID)) %>% 
  mutate(cluster_ID = factor(cluster_ID, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19"))) %>% 
  filter(mark == "hmC") %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  scale_fill_manual(values = c(Epi_colour)) +
  facet_grid(fragment ~ cluster_ID) +
  coord_cartesian(ylim = c(0,10)) +
  ggtitle("DNMT WT hmC") +
  theme(legend.position = "none", axis.title.x = element_blank(), plot.title = element_text(size = 14, hjust = 0.5))
```

Exporting

```{r warning=FALSE}
ggsave("../output/plots/ES_to_Epi/DNMT_ES_to_Epi_hmC.png", width = 20, height = 5, units = "in")
ggsave("../output/plots/ES_to_Epi/DNMT_ES_to_Epi_hmC.svg", width = 20, height = 5, units = "in")
```

Only one row for CHiC


```{r, fig.width=20, fig.height=3}
DNMT_WT %>% 
  mutate(cluster_ID = gsub("cluster", "", cluster_ID)) %>% 
  mutate(cluster_ID = factor(cluster_ID, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19"))) %>% 
  filter(mark == "CHiC") %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5) +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ cluster_ID) +
  coord_cartesian(ylim = c(0,10)) +
  ggtitle("DNMT WT CHiC") +
  theme(legend.position = "none", axis.title.x = element_blank(), plot.title = element_text(size = 14, hjust = 0.5))
```
Exporting

```{r}
ggsave("../output/plots/ES_to_Epi/DNMT_ES_to_Epi_CHiC.png", width = 20, height = 3, units = "in")
ggsave("../output/plots/ES_to_Epi/DNMT_ES_to_Epi_CHiC.svg", width = 20, height = 3, units = "in")
```



























