---
title: "Figure 3: enhancer marks in KOs"
output: html_notebook
---

This script will create the panels of Figure 3 of the manuscript using the EM and Louvain clusters which were identified during the paper revision process.

Defining source

```{r}
source("cluster_analysis_functions.R")
```
Importing data

```{r}
read_tsv("../data/interaction_pairs_uncapped.txt") -> magic_table_uncapped
```


## Marks at active enhancers

### EM clusters

#### Figure 3a

Filtering the magic table for relevant values and calculating KO to WT difference

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "active P - active E") %>% 
  select(interaction_ID, contains("EpiLC_D2"), -starts_with("P_"), -starts_with("CHiC"), -contains("mC_"), -contains("ATAC"),
         -contains("CTCF"), -contains("H3K27me3"), -contains("H3K4me3"), -contains("Input")) %>% 
  pivot_longer(2:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  mutate(sample = gsub("RNA", "PIR_RNA", sample)) %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-fragment, -stage, -day) %>% 
  pivot_wider(names_from = genotype, values_from = log2RPKM) %>% 
  mutate(KO_WT_diff = KO - WT)-> enhancer_marks_EM

head(enhancer_marks_EM)
```



Plotting Figure 3a (EM clusters)

```{r, fig.height=4.5, fig.width=5}
enhancer_marks_EM %>% 
  mutate(line = gsub("DNMT", "Dnmt TKO - WT", line)) %>% 
  mutate(line = gsub("TET", "Tet TKO - WT", line)) %>% 
  mutate(mark = gsub("RNA", "mRNA expression", mark)) %>% 
  ggplot(aes(KO_WT_diff, colour = mark)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_density() +
  xlim(-5, 5) +
  scale_colour_manual(values = c(K27ac_colour, K4me1_colour, exp_colour)) +
  xlab("Difference (log2 RPKM)") +
  ggtitle("EM clusters") +
  facet_wrap(facets = "line") +
  theme(legend.title = element_blank(), strip.background = element_rect(fill = "white"), legend.position = "inside", 
        legend.position.inside = c(0.755, 0.82), legend.text = element_text(size = 12), legend.box.background = element_rect(colour = "black"),
        panel.grid = element_blank(), plot.title = element_text(size = 14, hjust = 0.5)) -> figure3a_EM

figure3a_EM
```

### Louvain clusters

#### Figure 3a

Filtering the magic table for relevant values and calculating KO to WT difference

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_group == "active P - active E") %>% 
  select(interaction_ID, contains("EpiLC_D2"), -starts_with("P_"), -starts_with("CHiC"), -contains("mC_"), -contains("ATAC"),
         -contains("CTCF"), -contains("H3K27me3"), -contains("H3K4me3"), -contains("Input")) %>% 
  pivot_longer(2:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  mutate(sample = gsub("RNA", "PIR_RNA", sample)) %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-fragment, -stage, -day) %>% 
  pivot_wider(names_from = genotype, values_from = log2RPKM) %>% 
  mutate(KO_WT_diff = KO - WT)-> enhancer_marks_Louvain

head(enhancer_marks_Louvain)
```

Plotting Figure 3a (Louvain clusters)

```{r, fig.height=4.5, fig.width=5}
enhancer_marks_Louvain %>% 
  mutate(line = gsub("DNMT", "Dnmt TKO - WT", line)) %>% 
  mutate(line = gsub("TET", "Tet TKO - WT", line)) %>% 
  mutate(mark = gsub("RNA", "mRNA expression", mark)) %>% 
  ggplot(aes(KO_WT_diff, colour = mark)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_density() +
  xlim(-5, 5) +
  scale_colour_manual(values = c(K27ac_colour, K4me1_colour, exp_colour)) +
  xlab("Difference (log2 RPKM)") +
  ggtitle("Louvain clusters") +
  facet_wrap(facets = "line") +
  theme(legend.title = element_blank(), strip.background = element_rect(fill = "white"), legend.position = "inside", 
        legend.position.inside = c(0.755, 0.82), legend.text = element_text(size = 12), legend.box.background = element_rect(colour = "black"),
        panel.grid = element_blank(), plot.title = element_text(size = 14, hjust = 0.5)) -> figure3a_Louvain

figure3a_Louvain
```

Exporting plots

```{r}
ggsave("../output/plots/Figure3a_EM.svg",  figure3a_EM, width = 5, height = 4.5, units = "in")
ggsave("../output/plots/Figure3a_EM.png",  figure3a_EM, width = 5, height = 4.5, units = "in")
ggsave("../output/plots/Figure3a_Louvain.svg",  figure3a_Louvain, width = 5, height = 4.5, units = "in")
ggsave("../output/plots/Figure3a_Louvain.png",  figure3a_Louvain, width = 5, height = 4.5, units = "in")
```

Moving on to Figure 3c

## H3K4me1 marks at promoter - enhancer pairs

### EM clusters

### primed, active and poised enhancers

Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "inactive P - primed E" | EM_cluster_group == "active P - active E" | EM_cluster_group == "bivalent P - poised E") %>% 
  select(interaction_ID, EM_cluster_group, contains("H3K4me1")) %>% 
  select(interaction_ID, EM_cluster_group, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) -> K4me1_enhancers_EM

head(K4me1_enhancers_EM)
```

Plotting H3K4me1 for enhancer types - EM clusters

```{r, fig.width=3.5, fig.height=4}
K4me1_enhancers_EM %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "inactive P - primed E" ~ "Primed E",
                                      EM_cluster_group == "active P - active E" ~ "Active E",
                                      EM_cluster_group == "bivalent P - poised E" ~ "Poised E")) %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Primed E", "Active E", "Poised E"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("EM clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> K4me1_enhancer_violins_EM

K4me1_enhancer_violins_EM
```
### Louvain clusters

### primed, active and poised enhancers

Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_group == "inactive P - primed E" | Louvain_cluster_group == "active P - active E" | Louvain_cluster_group == "bivalent P - poised E") %>% 
  select(interaction_ID, Louvain_cluster_group, contains("H3K4me1")) %>% 
  select(interaction_ID, Louvain_cluster_group, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) -> K4me1_enhancers_Louvain

head(K4me1_enhancers_Louvain)
```
Plotting H3K4me1 for enhancer types - Louvain clusters

```{r, fig.width=3.5, fig.height=4}
K4me1_enhancers_Louvain %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "inactive P - primed E" ~ "Primed E",
                                      Louvain_cluster_group == "active P - active E" ~ "Active E",
                                      Louvain_cluster_group == "bivalent P - poised E" ~ "Poised E")) %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Primed E", "Active E", "Poised E"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("Louvain clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> K4me1_enhancer_violins_Louvain

K4me1_enhancer_violins_Louvain
```
Exporting H3K4me1 plots

```{r}
ggsave("../output/plots/Figure3c_EM.png", K4me1_enhancer_violins_EM, width = 3.5, height = 4, units = "in")
ggsave("../output/plots/Figure3c_EM.svg", K4me1_enhancer_violins_EM, width = 3.5, height = 4, units = "in")
ggsave("../output/plots/Figure3c_Louvain.png", K4me1_enhancer_violins_Louvain, width = 3.5, height = 4, units = "in")
ggsave("../output/plots/Figure3c_Louvain.svg", K4me1_enhancer_violins_Louvain, width = 3.5, height = 4, units = "in")
```

## H3K27ac marks at promoter - enhancer pairs

### EM clusters

### primed, active and poised enhancers

Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "inactive P - primed E" | EM_cluster_group == "active P - active E" | EM_cluster_group == "bivalent P - poised E") %>% 
  select(interaction_ID, EM_cluster_group, contains("H3K27ac")) %>% 
  select(interaction_ID, EM_cluster_group, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) -> K27ac_enhancers_EM

head(K27ac_enhancers_EM)
```

Plotting H3K27ac for enhancer types - EM clusters

```{r, fig.width=3.5, fig.height=4}
K27ac_enhancers_EM %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "inactive P - primed E" ~ "Primed E",
                                      EM_cluster_group == "active P - active E" ~ "Active E",
                                      EM_cluster_group == "bivalent P - poised E" ~ "Poised E")) %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Primed E", "Active E", "Poised E"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("EM clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> K27ac_enhancer_violins_EM

K27ac_enhancer_violins_EM
```

### Louvain clusters

### primed, active and poised enhancers

Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_group == "inactive P - primed E" | Louvain_cluster_group == "active P - active E" | Louvain_cluster_group == "bivalent P - poised E") %>% 
  select(interaction_ID, Louvain_cluster_group, contains("H3K27ac")) %>% 
  select(interaction_ID, Louvain_cluster_group, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) -> K27ac_enhancers_Louvain

head(K27ac_enhancers_Louvain)
```

Plotting H3K27ac for enhancer types - Louvain clusters

```{r, fig.width=3.5, fig.height=4}
K27ac_enhancers_Louvain %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "inactive P - primed E" ~ "Primed E",
                                      Louvain_cluster_group == "active P - active E" ~ "Active E",
                                      Louvain_cluster_group == "bivalent P - poised E" ~ "Poised E")) %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Primed E", "Active E", "Poised E"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("Louvain clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> K27ac_enhancer_violins_Louvain

K27ac_enhancer_violins_Louvain
```

Exporting H3K27ac plots

```{r}
ggsave("../output/plots/Figure3d_EM.png", K27ac_enhancer_violins_EM, width = 3.5, height = 4, units = "in")
ggsave("../output/plots/Figure3d_EM.svg", K27ac_enhancer_violins_EM, width = 3.5, height = 4, units = "in")
ggsave("../output/plots/Figure3d_Louvain.png", K27ac_enhancer_violins_Louvain, width = 3.5, height = 4, units = "in")
ggsave("../output/plots/Figure3d_Louvain.svg", K27ac_enhancer_violins_Louvain, width = 3.5, height = 4, units = "in")
```

## H3K27me3 marks at promoter - PIR pairs

### EM clusters

### polycomb promoters and poised enhancers

Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "polycomb P - unmarked PIR" | EM_cluster_group == "bivalent P - poised E") %>% 
  select(interaction_ID, EM_cluster_group, contains("H3K27me3")) %>% 
  select(interaction_ID, EM_cluster_group, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) -> K27me3_enhancers_EM

head(K27me3_enhancers_EM)
```

Plotting H3K27me3 for PIR types - EM clusters

```{r, fig.width=2.7, fig.height=4}
K27me3_enhancers_EM %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "polycomb P - unmarked PIR" ~ "Polycomb P",
                                      EM_cluster_group == "bivalent P - poised E" ~ "Poised E")) %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Polycomb P", "Poised E"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("EM clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> K27me3_enhancer_violins_EM

K27me3_enhancer_violins_EM
```
### Louvain clusters

### polycomb promoters and poised enhancers

Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_group == "polycomb P - unmarked PIR" | Louvain_cluster_group == "bivalent P - poised E") %>% 
  select(interaction_ID, Louvain_cluster_group, contains("H3K27me3")) %>% 
  select(interaction_ID, Louvain_cluster_group, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) -> K27me3_enhancers_Louvain

head(K27me3_enhancers_Louvain)
```

Plotting H3K27me3 for PIR types - Louvain clusters

```{r, fig.width=2.7, fig.height=4}
K27me3_enhancers_Louvain %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "polycomb P - unmarked PIR" ~ "Polycomb P",
                                      Louvain_cluster_group == "bivalent P - poised E" ~ "Poised E")) %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Polycomb P", "Poised E"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("Louvain clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> K27me3_enhancer_violins_Louvain

K27me3_enhancer_violins_Louvain
```

Exporting H3K27me3 plots

```{r}
ggsave("../output/plots/Figure3e_EM.png", K27me3_enhancer_violins_EM, width = 2.7, height = 4, units = "in")
ggsave("../output/plots/Figure3e_EM.svg", K27me3_enhancer_violins_EM, width = 2.7, height = 4, units = "in")
ggsave("../output/plots/Figure3e_Louvain.png", K27me3_enhancer_violins_Louvain, width = 2.7, height = 4, units = "in")
ggsave("../output/plots/Figure3e_Louvain.svg", K27me3_enhancer_violins_Louvain, width = 2.7, height = 4, units = "in")
```