---
title: "Figure 6bcde blood enhancers"
output:
  html_document:
    df_print: paged
---

The workbook will create Figure 6bcde relating to blood enhancers in EpiLCs. This uses the EM and Louvain clustering we have switched to during the paper revision.
I will attempt a different representation of the data as I find the original figure quite messy. Let's see how that goes.


Loading packages and options

```{r}
library(tidyverse)
options(scipen = 999)
source("cluster_analysis_functions.R")
```


Importing data

```{r}
read_tsv("../data/interaction_pairs_uncapped.txt") -> magic_table_uncapped
```

Filtering for the right marks and the right time points

```{r}
magic_table_uncapped %>% 
  select(-contains("cluster_ID"), -contains("UMAP"), -P_ID, -PIR_ID, -PIR_gene, -contains("_ES_"), - contains("_hmC_"), -contains("mC"),
         -contains("Input"), -contains("ATAC"), -contains("CTCF"), -contains("H3K27me3"), -contains("H3K27ac"), -contains("RNA"), -contains("P_"),
         -contains("H3K4me3"), P_gene) %>% 
  mutate(CHiC_diff_DNMT = CHiC_DNMT_KO_EpiLC_D2 - CHiC_DNMT_WT_EpiLC_D2,
         K4me1_diff_DNMT = PIR_H3K4me1_DNMT_KO_EpiLC_D2 - PIR_H3K4me1_DNMT_WT_EpiLC_D2,
         CHiC_diff_TET = CHiC_TET_KO_EpiLC_D2 - CHiC_TET_WT_EpiLC_D2,
         K4me1_diff_TET = PIR_H3K4me1_TET_KO_EpiLC_D2 - PIR_H3K4me1_TET_WT_EpiLC_D2) %>% 
  select(interaction_ID, EM_cluster_group, Louvain_cluster_group, P_gene, contains("diff")) -> mark_differences

head(mark_differences)
```

Defining genes of interest

```{r}
genes_of_interest <- c("Tal1", "Nfe2", "Runx1", "Klf1", "Lyl1", "Gata1")
```


### EM clustering

```{r}
mark_differences %>% 
  select(-Louvain_cluster_group) %>% 
  filter(P_gene %in% genes_of_interest) -> blood_PIR_differences_EM

head(blood_PIR_differences_EM)
```

Figure 6b EM (K4me1, TET)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  ggplot(aes(EM_cluster_group, K4me1_diff_TET, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference H3K4me1\n (TET TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6b_EM

Fig6b_EM
```

Figure 6c EM (CHiC, TET)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  ggplot(aes(EM_cluster_group, CHiC_diff_TET, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference CHiC\n (TET TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6c_EM

Fig6c_EM
```
Figure 6d EM (K4me1, DNMT)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  ggplot(aes(EM_cluster_group, K4me1_diff_DNMT, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference H3K4me1\n (DNMT TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6d_EM

Fig6d_EM
```

Figure 6e EM (CHiC, DNMT)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  filter (CHiC_diff_DNMT > -10) %>% 
  ggplot(aes(EM_cluster_group, CHiC_diff_DNMT, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference CHiC\n (DNMT TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6e_EM

Fig6e_EM
```
Exporting

```{r}
ggsave("../output/plots/Figure6b_K4me1_blood_enhancers_EM.png", Fig6b_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6b_K4me1_blood_enhancers_EM.svg", Fig6b_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6d_K4me1_blood_enhancers_EM.png", Fig6d_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6d_K4me1_blood_enhancers_EM.svg", Fig6d_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6c_CHiC_blood_enhancers_EM.png", Fig6c_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6c_CHiC_blood_enhancers_EM.svg", Fig6c_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6e_CHiC_blood_enhancers_EM.png", Fig6e_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6e_CHiC_blood_enhancers_EM.svg", Fig6e_EM, width = 5, height = 5, units = "in")
```

### Louvain clustering

```{r}
mark_differences %>% 
  select(-EM_cluster_group) %>% 
  filter(P_gene %in% genes_of_interest) -> blood_PIR_differences_Louvain

head(blood_PIR_differences_Louvain)
```

Figure 6b Louvain (K4me1, TET)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_Louvain %>% 
  ggplot(aes(Louvain_cluster_group, K4me1_diff_TET, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference H3K4me1\n (TET TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6b_Louvain

Fig6b_Louvain
```

Figure 6c Louvain (CHiC, TET)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_Louvain %>% 
  ggplot(aes(Louvain_cluster_group, CHiC_diff_TET, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference CHiC\n (TET TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6c_Louvain

Fig6c_Louvain
```

Figure 6d Louvain (K4me1, DNMT)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_Louvain %>% 
  ggplot(aes(Louvain_cluster_group, K4me1_diff_DNMT, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference H3K4me1\n (DNMT TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6d_Louvain

Fig6d_Louvain
```

Figure 6e Louvain (CHiC, DNMT)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_Louvain %>% 
  filter (CHiC_diff_DNMT > -10) %>% 
  ggplot(aes(Louvain_cluster_group, CHiC_diff_DNMT, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("difference CHiC\n (DNMT TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> Fig6e_Louvain

Fig6e_Louvain
```

Exporting

```{r}
ggsave("../output/plots/Figure6b_K4me1_blood_enhancers_Louvain.png", Fig6b_Louvain, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6b_K4me1_blood_enhancers_Louvain.svg", Fig6b_Louvain, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6d_K4me1_blood_enhancers_Louvain.png", Fig6d_Louvain, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6d_K4me1_blood_enhancers_Louvain.svg", Fig6d_Louvain, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6c_CHiC_blood_enhancers_Louvain.png", Fig6c_Louvain, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6c_CHiC_blood_enhancers_Louvain.svg", Fig6c_Louvain, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6e_CHiC_blood_enhancers_Louvain.png", Fig6e_Louvain, width = 5, height = 5, units = "in")
ggsave("../output/plots/Figure6e_CHiC_blood_enhancers_Louvain.svg", Fig6e_Louvain, width = 5, height = 5, units = "in")
```
