---
title: "Figure 4f - expression levels"
output: html_notebook
---

This workbook is creating the violin plots of RNA expression using in Figure 4f of stable and downregulated genes. This uses the EM and Louvain clustering we have switched to during the paper revision.

Starting with cluster group active P - active E (active promoter, active enhancer).
Having two groups of genes, the ones that change expression (ie in the TET TKO WT comparison) and those that don't change expression.


Loading packages and options

```{r}
library(tidyverse)
options(scipen = 999)
source("cluster_analysis_functions.R")
```


Importing data

```{r}
read_tsv("../data/interaction_pairs_uncapped.txt") -> magic_table_uncapped
read_tsv("../data/Annotated Probe Report for DE TET Epi WT vs KO no MTC.txt") -> TET_WT_vs_KO_import
```


Finding the group of downregulated genes (TET DE without MTC)

```{r}
TET_WT_vs_KO_import %>% 
  select(Probe, TET_WT_Epi_2, TET_KO_Epi_2) %>% 
  distinct(Probe, .keep_all = TRUE) %>% 
  filter(TET_KO_Epi_2 < TET_WT_Epi_2) %>% 
  pull(Probe) -> TET_KO_down_genes
```



## EM clustering

Plotting the expression data for TET KO down and stable genes that are found in the magic table.

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "active P - active E") %>% 
  select(interaction_ID, P_gene, RNA_TET_WT_EpiLC_D2, RNA_TET_KO_EpiLC_D2) %>% 
  drop_na() %>% 
  mutate(group = case_when(P_gene %in% TET_KO_down_genes ~ "TET_KO_down",
                           RNA_TET_WT_EpiLC_D2 - RNA_TET_KO_EpiLC_D2 > -0.25 & RNA_TET_WT_EpiLC_D2 - RNA_TET_KO_EpiLC_D2 < 0.25 ~ "stable_genes")) -> activeE_groups_EM

head(activeE_groups_EM)
```

How many do we get for downregulated and stable?

```{r}
activeE_groups_EM %>% 
  group_by(group) %>% 
  distinct(P_gene, .keep_all = TRUE) %>% # we need this here because we have several interactions per gene
  count() -> groups_counted_EM

groups_counted_EM

groups_counted_EM %>% 
  filter(group == "TET_KO_down") %>% 
  pull(n) -> number_KO_genes
```

I found that there are 405 interactions that involved a total of 145 TET KO down genes in this data frame and I want to have the same number of stable genes.

```{r}
set.seed(6)

activeE_groups_EM %>% 
  filter(group == "stable_genes") %>% 
  distinct(P_gene) %>% 
  slice_sample(n = number_KO_genes) %>% 
  pull() -> stable_genes_selected_EM
```

Now going back to the original magic table
filtering for the active P - active E cluster group (EM)
selecting expression
selecting stable and knocked down genes

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "active P - active E") %>% 
  select(interaction_ID, P_gene, RNA_TET_WT_EpiLC_D2, RNA_TET_KO_EpiLC_D2) %>% 
  mutate(group = case_when(P_gene %in% TET_KO_down_genes ~ "TET_KO_down",
                           P_gene %in% stable_genes_selected_EM ~ "stable")) %>% 
  filter(!is.na(group)) -> expression_data_EM

head(expression_data_EM)
```

Reshaping data for violins

```{r}
expression_data_EM %>% 
  rename(WT = RNA_TET_WT_EpiLC_D2, TKO = RNA_TET_KO_EpiLC_D2) %>% 
  pivot_longer(WT:TKO, names_to = "genotype", values_to = "log2RPKM") %>% 
  mutate(group = case_when(group == "TET_KO_down" ~ "Downregulated",
                          group == "stable" ~ "Stable")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "TKO"))) %>% 
  mutate(group = factor(group, levels = c("Stable", "Downregulated"))) -> expression_to_plot_EM

head(expression_to_plot_EM)
```
Plotting violins

```{r, fig.width=4, fig.height=4}
expression_to_plot_EM %>% 
  ggplot(aes(genotype, log2RPKM, fill = genotype)) +
  geom_violin(draw_quantiles = 0.5) +
  scale_fill_manual(values = c(TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  ylab("mRNA expression log2 RPKM") +
  facet_wrap(facets = "group") +
  theme(legend.position = "none", axis.title.x = element_blank(), strip.background = element_rect(fill = "white"),
        axis.title.y = element_text(size = 12)) -> expression_violins_EM

expression_violins_EM
```

Exporting

```{r}
ggsave("../output/plots/Figure4f_expression_violins_EM.png", expression_violins_EM, width = 4, height = 4, units = "in")
ggsave("../output/plots/Figure4f_expression_violins_EM.svg", expression_violins_EM, width = 4, height = 4, units = "in")
```

## Louvain clustering

Plotting the expression data for TET KO down and stable genes that are found in the magic table.

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_group == "active P - active E") %>% 
  select(interaction_ID, P_gene, RNA_TET_WT_EpiLC_D2, RNA_TET_KO_EpiLC_D2) %>% 
  drop_na() %>% 
  mutate(group = case_when(P_gene %in% TET_KO_down_genes ~ "TET_KO_down",
                           RNA_TET_WT_EpiLC_D2 - RNA_TET_KO_EpiLC_D2 > -0.25 & RNA_TET_WT_EpiLC_D2 - RNA_TET_KO_EpiLC_D2 < 0.25 ~ "stable_genes")) -> activeE_groups_Louvain

head(activeE_groups_Louvain)
```

How many do we get for downregulated and stable?

```{r}
activeE_groups_Louvain %>% 
  group_by(group) %>% 
  distinct(P_gene, .keep_all = TRUE) %>% # we need this here because we have several interactions per gene
  count() -> groups_counted_Louvain

groups_counted_Louvain

groups_counted_Louvain %>% 
  filter(group == "TET_KO_down") %>% 
  pull(n) -> number_KO_genes
```

I found that there are 170 TET KO down genes in this data frame and I want to have the same number of stable genes.

```{r}
set.seed(6)

activeE_groups_Louvain %>% 
  filter(group == "stable_genes") %>% 
  distinct(P_gene) %>% 
  slice_sample(n = number_KO_genes) %>% 
  pull() -> stable_genes_selected_Louvain
```

Now going back to the original magic table
filtering for the active P - active E cluster group (Louvain)
selecting expression
selecting stable and knocked down genes

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_group == "active P - active E") %>% 
  select(interaction_ID, P_gene, RNA_TET_WT_EpiLC_D2, RNA_TET_KO_EpiLC_D2) %>% 
  mutate(group = case_when(P_gene %in% TET_KO_down_genes ~ "TET_KO_down",
                           P_gene %in% stable_genes_selected_Louvain ~ "stable")) %>% 
  filter(!is.na(group)) -> expression_data_Louvain

head(expression_data_Louvain)
```

Reshaping data for violins

```{r}
expression_data_Louvain %>% 
  rename(WT = RNA_TET_WT_EpiLC_D2, TKO = RNA_TET_KO_EpiLC_D2) %>% 
  pivot_longer(WT:TKO, names_to = "genotype", values_to = "log2RPKM") %>% 
  mutate(group = case_when(group == "TET_KO_down" ~ "Downregulated",
                          group == "stable" ~ "Stable")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "TKO"))) %>% 
  mutate(group = factor(group, levels = c("Stable", "Downregulated"))) -> expression_to_plot_Louvain

head(expression_to_plot_Louvain)
```

Plotting violins

```{r, fig.width=4, fig.height=4}
expression_to_plot_Louvain %>% 
  ggplot(aes(genotype, log2RPKM, fill = genotype)) +
  geom_violin(draw_quantiles = 0.5) +
  scale_fill_manual(values = c(TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  ylab("mRNA expression log2 RPKM") +
  facet_wrap(facets = "group") +
  theme(legend.position = "none", axis.title.x = element_blank(), strip.background = element_rect(fill = "white"),
        axis.title.y = element_text(size = 12)) -> expression_violins_Louvain

expression_violins_Louvain
```

Exporting

```{r}
ggsave("../output/plots/Figure4f_expression_violins_Louvain.png", expression_violins_Louvain, width = 4, height = 4, units = "in")
ggsave("../output/plots/Figure4f_expression_violins_Louvain.svg", expression_violins_Louvain, width = 4, height = 4, units = "in")
```
