---
title: "Enhancer redundancy - mark differences"
output:
  html_document:
    df_print: paged
---

This workbook is creating the heatmaps of ATAC, H3K27ac and H3K4me1 differences for active enhancers looking at redundancy of elements. This uses the EM and Louvain clustering we have switched to during the paper revision.

This script creates the heatmaps for Figure 4g and Supplemental Figure 6c and d in the revised manuscript.

Starting with cluster group active P - active E (active promoter, active enhancer).
Having two groups of genes, the ones that change expression (ie in the TET TKO WT comparison) and those that don't change expression.


Loading packages and options

```{r}
library(tidyverse)
library(pheatmap)
library(RColorBrewer)
options(scipen = 999)
source("cluster_analysis_functions.R")
```

Importing data

```{r}
read_tsv("../data/interaction_pairs_uncapped.txt") -> magic_table_uncapped
read_tsv("../data/Annotated Probe Report for DE TET Epi WT vs KO no MTC.txt") -> TET_WT_vs_KO_import
```

Finding the group of downregulated genes (TET DE without MTC)

```{r}
TET_WT_vs_KO_import %>% 
  select(Probe, TET_WT_Epi_2, TET_KO_Epi_2) %>% 
  distinct(Probe, .keep_all = TRUE) %>% 
  filter(TET_KO_Epi_2 < TET_WT_Epi_2) %>% 
  pull(Probe) -> TET_KO_down_genes
```



## EM clustering

Plotting the expression data for TET KO down and stable genes that are found in the magic table.

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "active P - active E") %>% 
  select(interaction_ID, P_gene, RNA_TET_WT_EpiLC_D2, RNA_TET_KO_EpiLC_D2) %>% 
  drop_na() %>% 
  mutate(group = case_when(P_gene %in% TET_KO_down_genes ~ "TET_KO_down",
                           RNA_TET_WT_EpiLC_D2 - RNA_TET_KO_EpiLC_D2 > -0.25 & RNA_TET_WT_EpiLC_D2 - RNA_TET_KO_EpiLC_D2 < 0.25 ~ "stable_genes")) -> activeE_groups_EM

head(activeE_groups_EM)
```
How many do we get for downregulated and stable?

```{r}
activeE_groups_EM %>% 
  group_by(group) %>% 
  distinct(P_gene, .keep_all = TRUE) %>% # we need this here because we have several interactions per gene
  count() -> groups_counted_EM

groups_counted_EM

groups_counted_EM %>% 
  filter(group == "TET_KO_down") %>% 
  pull(n) -> number_KO_genes
```


I found that there are 405 interactions that involved a total of 145 TET KO down genes in this data frame and I want to have the same number of stable genes.

```{r}
set.seed(6)

activeE_groups_EM %>% 
  filter(group == "stable_genes") %>% 
  distinct(P_gene) %>% 
  slice_sample(n = number_KO_genes) %>% 
  pull() -> stable_genes_selected_EM
```

Now going back to the original magic table
filtering for the active P - active E cluster group (EM)
selecting relevant marks
selecting stable and knocked down genes

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "active P - active E") %>% 
  select(interaction_ID, P_gene, RNA_TET_WT_EpiLC_D2, RNA_TET_KO_EpiLC_D2, PIR_ATAC_TET_WT_EpiLC_D2, PIR_ATAC_TET_KO_EpiLC_D2, PIR_H3K27ac_TET_WT_EpiLC_D2,
         PIR_H3K27ac_TET_KO_EpiLC_D2, PIR_H3K4me1_TET_WT_EpiLC_D2,
         PIR_H3K4me1_TET_KO_EpiLC_D2) %>% 
  mutate(group = case_when(P_gene %in% TET_KO_down_genes ~ "TET_KO_down",
                           P_gene %in% stable_genes_selected_EM ~ "stable")) %>% 
  filter(!is.na(group)) -> redundancy_plot_data_EM

head(redundancy_plot_data_EM)
```



I need to substantially alter the shape of this dataframe. I want each enhancer per gene to be one column, and fill up the empty columns with NAs.

### ATAC

```{r}
redundancy_plot_data_EM %>% 
  select(P_gene, group, PIR_ATAC_TET_WT_EpiLC_D2) %>% 
  group_by(P_gene) %>% 
  arrange(desc(PIR_ATAC_TET_WT_EpiLC_D2), .by_group = TRUE) %>% 
  mutate(enhancer = row_number()) %>% 
  add_count() %>% 
  pivot_wider(names_from = enhancer, values_from = PIR_ATAC_TET_WT_EpiLC_D2, values_fill = NA) -> ATAC_for_hm_EM

head(ATAC_for_hm_EM)
```

Preparing data the heatmap for stable (EM)

```{r}
ATAC_for_hm_EM %>% 
  filter(group == "stable") %>% 
  arrange(desc(n)) %>% 
  select(-n) %>% 
  select(P_gene, 3:12) %>% 
  as.data.frame() %>% 
  column_to_rownames("P_gene") -> heatmap.stable.ATAC.EM

head(heatmap.stable.ATAC.EM)
```


Preparing data the heatmap for TET KO down (EM)

```{r}
ATAC_for_hm_EM %>% 
  filter(group == "TET_KO_down") %>% 
  arrange(desc(n)) %>% 
  select(-n) %>% 
  select(P_gene, 3:12) %>% 
  as.data.frame() %>% 
  column_to_rownames("P_gene") -> heatmap.down.ATAC.EM

head(heatmap.down.ATAC.EM)
```
Drawing the heatmap for stable

```{r}
pheatmap(heatmap.stable.ATAC.EM,
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         #breaks = seq(0.5, 1.5, length.out = 100),
         color = colorRampPalette(brewer.pal(3, "YlOrRd"))(99),
         border_color = "#d4d4d4",
         na_col = "#f5f5f5",
         angle_col = 0,
         main = "ATAC TET WT EPI - stable genes",
         show_rownames = FALSE) -> heatmap.stable.export.ATAC.EM

ggsave("../output/plots/SuppFigure6b_stable_EM.png", heatmap.stable.export.ATAC.EM, width = 5, height = 7, units = "in")
ggsave("../output/plots/SuppFigure6b_stable_EM.svg", heatmap.stable.export.ATAC.EM, width = 5, height = 7, units = "in")
```


Drawing the heatmap for TET KO down

```{r}
pheatmap(heatmap.down.ATAC.EM,
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         #breaks = seq(0.5, 1.5, length.out = 100),
         color = colorRampPalette(brewer.pal(3, "YlOrRd"))(99),
         border_color = "#d4d4d4",
         na_col = "#f5f5f5",
         angle_col = 0,
         main = "ATAC TET WT EPI - TET KO downregulated",
         show_rownames = FALSE
         ) -> heatmap.down.export

ggsave("../output/plots/SuppFigure6b_down_EM.png", heatmap.down.export, width = 5, height = 7, units = "in")
ggsave("../output/plots/SuppFigure6b_down_EM.svg", heatmap.down.export, width = 5, height = 7, units = "in")
```

