---
title: "Grouping clusters"
output: html_notebook
---

04/01/2024

I have recently reclustered our pairwise interaction data
1) EM clustering on umap dimensionality reduced data to 15 dimensions (previously wrong with 30 dimensions)
2) Louvain clustering after PCA dimensionality reduced data to 12 dimensions (as alternative clustering method).

When these clusters were grouped to the larger categories that we had previously defined on the TET WT Epi data there seemed to be a very reasonable overlap between the two approaches. However, this was inconveniently spread over two scripts, so here I'm collating the information.

In order to group the clusters I previously determined that it makes sense to use the capped data - otherwise the outlier clusters dominate the boxplots and most things look unmarked.

Libraries and options

```{r}
library(tidyverse)
options(scipen = 999)
```

Importing capped data and annotations

```{r}
read_tsv("../data/umap_mark_data_capped.txt") -> marks_capped
read_tsv("../data/EM_cluster_assignments.txt") -> EM_clusters_import
read_tsv("../data/Louvain_cluster_assignment.tsv") -> Louvain_clusters
```
Tidying and merging

```{r}
EM_clusters_import %>% 
  select(interaction_ID, cluster_ID_15dim) -> EM_clusters

marks_capped %>% 
  select(-cluster_ID, -cluster_group) %>% 
  left_join(EM_clusters) %>% 
  left_join(Louvain_clusters) %>% 
  relocate(cluster_ID_15dim, .after = interaction_ID) %>% 
  relocate(Louvain_cluster_ID, .after = cluster_ID_15dim) %>% 
  rename(EM_cluster_ID = cluster_ID_15dim) -> marks

head(marks)
```

Displaying the individual EM clusters

```{r, fig.width=4, fig.height=4}
marks %>% 
  sample_n(size = 7000) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = as.factor(EM_cluster_ID))) +
  geom_point(size = 0.5) + 
  scale_colour_viridis_d(option  = "turbo") +
  xlab("UMAP1") +
  ylab("UMAP2") +
  xlim(-10,12) +
  ylim(-10,7) +
  theme_bw(base_size = 16) +
  theme(aspect.ratio = 1, legend.title = element_blank(), legend.text = element_text(size = 6), axis.title = element_text(size = 12)) -> umap_plot_clusters_15dim

umap_plot_clusters_15dim


#ggsave(filename = "../output/plots/umap_plot_individual_clusters_EM.png", umap_plot_clusters_15dim, width = 4, height = 4, units = "in", device = "png")
#ggsave(filename = "../output/plots/umap_plot_individual_clusters_EM.svg", umap_plot_clusters_15dim, width = 4, height = 4, units = "in", device = "svg")
```

Displaying the individual Louvain clusters

Clusters to exclude (the ones with fewer than 1000 data points)

```{r}
excluded_clusters <- c("21", "22", "23", "24", "25", "26", "27", "28", "29", "30")
```

```{r, fig.height=4, fig.width=5}
marks %>% 
  #filter(!Louvain_cluster_ID %in% excluded_clusters) %>% 
  sample_n(size = 7000) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = as.factor(Louvain_cluster_ID))) +
  geom_point(size = 0.5) + 
  #geom_point(data = marks %>% filter(Louvain_cluster_ID == 30), colour = "red", size = 0.5) +
  scale_colour_viridis_d(option  = "turbo") +
  xlab("UMAP1") +
  ylab("UMAP2") +
  xlim(-10,12) +
  ylim(-10,7) +
  theme_bw(base_size = 16) +
  theme(aspect.ratio = 1, legend.title = element_blank(), legend.text = element_text(size = 10), axis.title = element_text(size = 12)) -> umap_plot_clusters_Lou

umap_plot_clusters_Lou


#ggsave(filename = "../output/plots/umap_plot_individual_clusters_Lou.png", umap_plot_clusters_Lou, width = 6, height = 4, units = "in", device = "png")
#ggsave(filename = "../output/plots/umap_plot_individual_clusters_Lou.svg", umap_plot_clusters_Lou, width = 6, height = 4, units = "in", device = "png")
```

Let's plot the marks for the clusters. For this, we'll only consider TET WT Epi so I'll make a smaller dataframe.

```{r}
marks %>% 
  select(-contains("Input"), -contains("KO_"), -contains("ES"), -contains("DNMT")) -> marks_TET_WT_Epi

head(marks_TET_WT_Epi)

```

I'm currently not adding the RNA data which makes the object much smaller. A normal pivot should work as this will need less memory than in for the big table I used in the other script.

```{r}
marks_TET_WT_Epi %>% 
  select(-UMAP1, -UMAP2, -P_ID, -PIR_ID, -P_gene, -PIR_gene, -contains("CHiC")) %>% 
  pivot_longer(P_hmC_TET_WT_EpiLC_D2:last_col(), names_to = "samples") %>% 
  separate(samples, into = c("fragment", "mark"), extra = "drop") %>% 
  mutate(EM_cluster_ID = factor(EM_cluster_ID, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19"))) %>% 
  mutate(Louvain_cluster_ID = factor(Louvain_cluster_ID, levels = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20", "21","22","23","24","25","26","27","28","29","30"))) -> marks_TET_WT_Epi_long

head(marks_TET_WT_Epi_long)
```
Plotting EM cluster marks

```{r, fig.width=8, fig.height=10}
marks_TET_WT_Epi_long %>% 
  filter(fragment == "P") %>% 
  ggplot(aes(EM_cluster_ID, value, fill = EM_cluster_ID)) +
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(facets = "mark", scale = "free", ncol = 2) +
  scale_fill_viridis_d(option = "turbo") +
  ggtitle("P TET WT Epi marks by EM cluster") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(size = 10, angle = 90, vjust = 0.3), plot.title = element_text(size = 14, hjust = 0.5), legend.position = "none") -> P_marks_plot_EM

P_marks_plot_EM
```

```{r, fig.width=8, fig.height=10}
marks_TET_WT_Epi_long %>% 
  filter(fragment == "PIR") %>% 
  ggplot(aes(EM_cluster_ID, value, fill = EM_cluster_ID)) +
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(facets = "mark", scale = "free", ncol = 2) +
  scale_fill_viridis_d(option = "turbo") +
  ggtitle("PIR TET WT Epi marks by EM cluster") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(size = 10, angle = 90, vjust = 0.3), plot.title = element_text(size = 14, hjust = 0.5), legend.position = "none") -> PIR_marks_plot_EM

PIR_marks_plot_EM
```

Plotting Louvain cluster marks

```{r, fig.width=8, fig.height=10}
marks_TET_WT_Epi_long %>% 
  filter(fragment == "P") %>% 
  filter(!Louvain_cluster_ID %in% excluded_clusters) %>% 
  ggplot(aes(Louvain_cluster_ID, value, fill = Louvain_cluster_ID)) +
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(facets = "mark", scale = "free", ncol = 2) +
  scale_fill_viridis_d(option = "turbo") +
  ggtitle("P TET WT Epi marks by Louvain cluster") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(size = 10, angle = 90, vjust = 0.3), plot.title = element_text(size = 14, hjust = 0.5), legend.position = "none") -> P_marks_plot_Louvain

P_marks_plot_Louvain
```

```{r, fig.width=8, fig.height=10}
marks_TET_WT_Epi_long %>% 
  filter(fragment == "PIR") %>% 
  filter(!Louvain_cluster_ID %in% excluded_clusters) %>% 
  ggplot(aes(Louvain_cluster_ID, value, fill = Louvain_cluster_ID)) +
  geom_boxplot(outlier.size = 0.5) +
  facet_wrap(facets = "mark", scale = "free", ncol = 2) +
  scale_fill_viridis_d(option = "turbo") +
  ggtitle("PIR TET WT Epi marks by Louvain cluster") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(size = 10, angle = 90, vjust = 0.3), plot.title = element_text(size = 14, hjust = 0.5), legend.position = "none") -> PIR_marks_plot_Louvain

PIR_marks_plot_Louvain
```

Exporting

```{r}
ggsave(filename = "../output/plots/P_marks_plot_EM.png", P_marks_plot_EM, width = 8, height = 10, units = "in", device = "png")
ggsave(filename = "../output/plots/PIR_marks_plot_EM.png", PIR_marks_plot_EM, width = 8, height = 10, units = "in", device = "png")
ggsave(filename = "../output/plots/P_marks_plot_Louvain.png", P_marks_plot_Louvain, width = 8, height = 10, units = "in", device = "png")
ggsave(filename = "../output/plots/PIR_marks_plot_Louvain.png", PIR_marks_plot_Louvain, width = 8, height = 10, units = "in", device = "png")
```

Soo .... I stared long and hard at these boxplots (again) and think we should make the following groupings. It's quite nice to see that EM and Louvain clustering come out relatively similar.

EM cluster groupings

```{r}
EM_activeP_lowmarkedPIR <- c("4", "7", "10", "11", "12", "13", "18", "16")
EM_pcP_unmarkedPIR <- c("0", "2", "9")
EM_inactiveP_unmarkedPIR <- c("1", "5", "6", "15")
EM_inactiveP_CTCFprimedE <- c("19")
EM_inactiveP_primedE <- c("14")
EM_bivalentP_poisedE <- c( "3")
EM_activeP_activeP <- c("17")
EM_activeP_activeE <- c("8")
```


Louvain cluster grouping

```{r}
Louvain_activeP_lowmarkedPIR <- c("2", "5", "9", "13", "15", "16")
Louvain_pcP_unmarkedPIR <- c("0", "12", "17", "30")
Louvain_inactiveP_unmarkedPIR <- c("3", "4", "7", "9", "10", "14", "21", "22", "23", "26", "27", "28", "29")
Louvain_inactiveP_CTCFprimedE <- c("1", "19")
Louvain_inactiveP_primedE <- c("20", "6")
Louvain_bivalentP_poisedE <- c( "18", "25")
Louvain_activeP_activeP <- c("11", "24")
Louvain_activeP_activeE <- c("8")
```

In the end I assigned the previously excluded clusters to the groupings based on their location on the UMAP.

Introducing the cluster groups

```{r}
marks %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_ID %in% Louvain_activeP_lowmarkedPIR ~ "active P - low marked PIR",
                                           Louvain_cluster_ID %in% Louvain_pcP_unmarkedPIR ~ "polycomb P - unmarked PIR",
                                           Louvain_cluster_ID %in% Louvain_inactiveP_unmarkedPIR ~ "inactive P - unmarked PIR",
                                           Louvain_cluster_ID %in% Louvain_inactiveP_CTCFprimedE ~ "inactive P - primed E, CTCF",
                                           Louvain_cluster_ID %in% Louvain_inactiveP_primedE ~ "inactive P - primed E",
                                           Louvain_cluster_ID %in% Louvain_bivalentP_poisedE ~ "bivalent P - poised E",
                                           Louvain_cluster_ID %in% Louvain_activeP_activeP ~ "active P - active P",
                                           Louvain_cluster_ID %in% Louvain_activeP_activeE ~ "active P - active E")) %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_ID %in% EM_activeP_lowmarkedPIR ~ "active P - low marked PIR",
                                           EM_cluster_ID %in% EM_pcP_unmarkedPIR ~ "polycomb P - unmarked PIR",
                                           EM_cluster_ID %in% EM_inactiveP_unmarkedPIR ~ "inactive P - unmarked PIR",
                                           EM_cluster_ID %in% EM_inactiveP_CTCFprimedE ~ "inactive P - primed E, CTCF",
                                           EM_cluster_ID %in% EM_inactiveP_primedE ~ "inactive P - primed E",
                                           EM_cluster_ID %in% EM_bivalentP_poisedE ~ "bivalent P - poised E",
                                           EM_cluster_ID %in% EM_activeP_activeP ~ "active P - active P",
                                           EM_cluster_ID %in% EM_activeP_activeE ~ "active P - active E")) %>%
  relocate(Louvain_cluster_group, .after = Louvain_cluster_ID) %>% 
  relocate(EM_cluster_group, .after = Louvain_cluster_ID) -> marks_groups

head(marks_groups)
```
Some stats

How many excluded Louvain clusters (fewer than 1000 data points)
```{r}
marks_groups %>% 
  filter(is.na(Louvain_cluster_group)) %>% 
  count()
```
Losing about 1 %. (This is with the small clusters excluded.)
However, later on I assigned them to a group, therefore there are now none lost.

How ofter is the cluster group the same between Louvain and EM?

```{r}
marks_groups %>% 
  filter(EM_cluster_group == Louvain_cluster_group) %>% 
  count() -> same

same/nrow(marks_groups)*100
```
70 % overlap. Not as much as I would have expected, but not bad. It's going to be mainly in the continuum of the markings.

Let's look at this in a bit more detail

```{r}
marks_groups %>% 
  select(interaction_ID, EM_cluster_group, Louvain_cluster_group) %>% 
  drop_na() %>% 
  mutate(same = if_else(EM_cluster_group == Louvain_cluster_group, TRUE, FALSE, missing = FALSE)) %>% 
  group_by(EM_cluster_group, same) %>% 
  tally() %>% 
  pivot_wider(names_from = same, values_from = n) %>% 
  mutate(total = `FALSE` + `TRUE`) %>% 
  mutate(percent_found_in_Louvain = `TRUE`/total*100)
  
```

The primed enhancers are a bit low for my taste ...

Let's visualise

Defining the colours

```{r}
cluster_group_names <- c("active P - low marked PIR",
                        "polycomb P - unmarked PIR",
                        "inactive P - unmarked PIR",
                        "inactive P - primed E, CTCF",
                        "inactive P - primed E",
                        "bivalent P - poised E",
                        "active P - active P",
                        "active P - active E"
                        )
cluster_group_colours <- c("#b785bb", "#bf6c6c", "#b2b5b5", "#741113", "#eaca48", "#ca3526", "#91c740", "#252d37")
names(cluster_group_colours) <- cluster_group_names
```

Plotting EM clusters

```{r, fig.width=8, fig.height=4}
marks_groups %>% 
  slice_sample(prop = 0.04) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = EM_cluster_group)) +
  geom_point(size = 0.3) +
  theme_bw(base_size = 14) +
  ggtitle("EM clusters") +
  coord_cartesian(xlim = c(-10, 12), ylim = c(-10, 8)) +
  scale_colour_manual(values = cluster_group_colours) +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5), legend.title = element_blank()) -> UMAP_EM_clusters

UMAP_EM_clusters
```
  
Exporting

```{r}
ggsave("../output/plots/UMAP_EM_clusters.png", UMAP_EM_clusters, width = 8, height = 4, unit = "in")
ggsave("../output/plots/UMAP_EM_clusters.svg", UMAP_EM_clusters, width = 8, height = 4, unit = "in")
```



Plotting Louvain clusters

```{r, fig.width=8, fig.height=4}
marks_groups %>% 
  slice_sample(prop = 0.04) %>% 
  ggplot(aes(UMAP1, UMAP2, colour = Louvain_cluster_group)) +
  geom_point(size = 0.3) +
  theme_bw(base_size = 14) +
  ggtitle("Louvain clusters") +
  coord_cartesian(xlim = c(-10, 12), ylim = c(-10, 8)) +
  scale_colour_manual(values = cluster_group_colours) +
  theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5), legend.title = element_blank()) -> UMAP_Louvain_clusters

UMAP_Louvain_clusters
```
  
Exporting

```{r}
ggsave("../output/plots/UMAP_Louvain_clusters.png", UMAP_Louvain_clusters, width = 8, height = 4, unit = "in")
ggsave("../output/plots/UMAP_Louvain_clusters.svg", UMAP_Louvain_clusters, width = 8, height = 4, unit = "in")
```




Let's create the new 'magic tables' with these groups

Importing and tidying the expression data

```{r}
read_tsv("../data/pluripotency_exit_RNA_Epi_and_EB_log2RPKM.txt") -> expression_import

expression_import %>% 
  select(Probe, DNMT_WT_ES:last_col()) %>% 
  rename_with(~gsub("_ES", "_ES_0",.)) %>% 
  rename_with(~paste0("RNA_", .)) %>% 
  rename(Probe = RNA_Probe) %>% 
  distinct(Probe, .keep_all = TRUE) -> expression

head(expression)
```
Adding the expression data to the capped marks data

```{r}
marks_groups %>% 
  left_join(expression, by = c("P_gene" = "Probe")) -> interaction_pair_attributes_capped
```

Now replacing the capped values with the uncapped ones

```{r}
read_tsv("../data/umap_mark_data_uncapped.txt") -> uncapped_import

interaction_pair_attributes_capped %>% 
  select(interaction_ID:Louvain_cluster_group) %>% 
  left_join(uncapped_import) %>% 
  left_join(expression, by = c("P_gene" = "Probe")) -> interaction_pair_attributes_uncapped
```
Exporting the new 'magic tables'

```{r}
write_tsv(interaction_pair_attributes_capped, "../output/tables/interaction_pairs_capped.txt")
write_tsv(interaction_pair_attributes_uncapped, "../output/tables/interaction_pairs_uncapped.txt")
```

