---
title: "Methylation of P/PIR of DNMT KO upregulated genes"
output: html_notebook
---


Reviewer's comment:

For example, the authors mention that “Dnmt TKO resulted in more upregulated than downregulated genes in EpiLC, consistent with a role in gene repression”. How many of the upregulated genes are associated with regulatory elements (promoter or enhancers) that actually get methylated during wildtype ESC-EpiLC transition, thereby implying direct responses in mutants. 

- Identify P/PIR of DNMT KO upregulated genes
- plot methylation DNMT ES WT to DNMT Epi WT

Defining source, libraries and options

```{r}
source("cluster_analysis_functions.R")
options(scipen = 0)
```

Importing datasets

The marks table was created for capped and uncapped values (the latter only to be used for colouring the umaps).
Linkings of marks to clusters happens in the script series umap_20_01_20_unccapped_marks and was exported from there.

```{r}
#read_tsv("../data/clusters_20_capped_marks_220517.txt") -> clusters_capped_marks
read_tsv("../data/clusters_20_uncapped_marks_220517.txt") -> clusters_uncapped_marks
#read_tsv("../data/pluripotency_exit_RNA_Epi_and_EB_log2RPKM.txt") -> expression_import
read_tsv("../data/Annotated Probe Report for DE DNMT Epi WT vs KO no MTC.txt") -> DNMT_KO_Epi_upregulated_import
```

```{r}
clusters_uncapped_marks %>% 
  rename_with(~gsub("H3K27Ac", "H3K27ac",.)) -> marks
colnames(marks)
```
Getting the gene names for the DNMT KO Epi upregulated genes

```{r}
DNMT_KO_Epi_upregulated_import %>% 
  pull(Probe) -> DNMT_KO_Epi_up_genes
```

Filtering the marks table for methylation of elements associated with the upregulated genes

```{r}
marks %>% 
  filter(bait_gene %in% DNMT_KO_Epi_up_genes) %>% 
  select(interaction_ID, bait_ID, oe_ID, bait_gene, cluster_ID, contains("mC"), -contains("hmC"), -contains("TET")) -> methylation_up_genes

head(methylation_up_genes)
```
Since we are really only looking at clusters 2, 12, 16, 17, maybe we should filter to only include those clusters in the analysis

```{r}
included_clusters <- c("cluster2", "cluster12", "cluster16", "cluster17")

methylation_up_genes %>% 
  filter(cluster_ID %in% included_clusters) %>% 
  select(-cluster_ID) %>% 
  pivot_longer(5:last_col(), names_to = "condition", values_to = "methylation") %>% 
  separate(condition, into = c("fragment", "mC", "DNMT", "genotype", "stage")) %>% 
  select(-mC, -DNMT) %>% 
  mutate(stage = factor(stage, levels = c("ES", "Epi"))) %>% 
  mutate(fragment = gsub("bait", "P", fragment)) %>% 
  mutate(fragment = gsub("oe", "PIR", fragment)) -> methylation_included_clusters

head(methylation_included_clusters)

methylation_included_clusters %>% 
  filter(fragment == "P") %>% 
  distinct(bait_ID, genotype, stage, .keep_all = TRUE) -> deduplicated_promoters

methylation_included_clusters %>% 
  filter(fragment == "PIR") %>% 
  distinct(oe_ID, genotype, stage, .keep_all = TRUE) -> deduplicated_PIRs

deduplicated_promoters %>% 
  bind_rows(deduplicated_PIRs) -> methylation_included_clusters_dedup
```

Plotting methylation across fragment, genotype and stage

```{r, fig.width=4, fig.height=3}
methylation_included_clusters_dedup %>% 
  ggplot(aes(stage, methylation, fill = stage)) + 
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(cols = vars(fragment)) +
  ylab("% methylation") +
  theme(legend.position = "none", axis.title.x = element_blank()) -> methylation_upregulated_elements_violins


  methylation_upregulated_elements_violins
```

Exporting

```{r}
ggsave("../output/plots/methylation_DNMT_KO_up_genes_violins.png", methylation_upregulated_elements_violins, width = 4, height = 3)
ggsave("../output/plots/methylation_DNMT_KO_up_genes_violins.svg", methylation_upregulated_elements_violins, width = 4, height = 3)
```


Note by: I haven't deduplicated the fragments, so if a promoter appears in more than one interaction, it will be plotted several times. I reasoned that this is ok as it represents different PIRs.

Later note: Actually, it does make a difference. I notices that when plotting expression. I have therefore now deduplicated the both P and PIRs. For promoters it makes a difference (the violin is much smoother), for the PIRs it doesn't.
