---
title: "Re-clustering after UWOT"
output: html_notebook
---

When looking for alternative ways of clustering and explored the Louvain algoriths, I noticed that the original EM clustering had used a UMAP representation of 30 dimensions of data. However, there are only 17 variables for TET WT Epi - not all 50 plus for all the conditions. I need to double check that this is actually the case but I'm pretty sure. As this makes no sense, I need to redo the clustering. 

Packages and settings

```{r}
library(uwot)
library(tidyverse)
options(scipen = 999)
```


Importing the data

```{r}
read_tsv("../data/umap_mark_data_uncapped.txt") -> marks_import
head(marks_import)
colnames(marks_import)
```

Converting into a dataframe that only had the TET WT Epi data in

```{r}
marks_import %>% 
  select(interaction_ID, contains("TET")) %>% 
  select(interaction_ID, contains("EpiLC")) %>% 
  select(interaction_ID, contains("WT")) %>% 
  select(interaction_ID, contains("D2")) %>% 
  select(-contains("Input")) %>% 
  rename_with(~ gsub("WT_EpiLC_D2", "",.)) %>% 
  column_to_rownames(var = "interaction_ID") -> marks_for_umap

head(marks_for_umap)
```

Running umap

```{r}
uwot::umap(marks_for_umap, n_neighbors = 20, min_dist = 0.01, scale = TRUE) -> umap_2_dim

umap_2_dim %>% 
  data.frame() %>% 
  rownames_to_column(var = "interaction_ID") %>% 
  as_tibble() -> umap_2

head(umap_2)
```

Checking the output

```{r}
umap_2 %>% 
  sample_frac(size = 0.2) %>% 
  ggplot(aes(X1, X2)) +
  geom_point(size = 0.1) +
  coord_cartesian(xlim = c(-10, 10), ylim = c(-10, 10)) +
  theme(aspect.ratio = 1)
```

Dimensionality reduction for clustering

```{r}
uwot::umap(marks_for_umap, n_neighbors = 25, min_dist = 0.01, scale = TRUE, n_components = 15) -> umap_15_dim

umap_15_dim %>% 
  data.frame() %>% 
  rownames_to_column(var = "interaction_ID") %>% 
  as_tibble() -> umap_15

head(umap_15)
```
Exporting

```{r}
write_csv(umap_15, "../output/tables/dim_reduction_15_for_clustering.csv")
```



