---
title: "Supplemental Figure 8 defg blood enhancers"
output:
  html_document:
    df_print: paged
---

The workbook will create Supplemental Figure 8defg relating to blood enhancers in EpiLCs. This uses the EM and Louvain clustering we have switched to during the paper revision.
Same representation as in Figure 6


Loading packages and options

```{r}
library(tidyverse)
options(scipen = 999)
source("cluster_analysis_functions.R")
```


Importing data

```{r}
read_tsv("../data/interaction_pairs_uncapped.txt") -> magic_table_uncapped
```

Filtering for the right marks and the right time points

```{r}
magic_table_uncapped %>% 
  select(-contains("cluster_ID"), -contains("UMAP"), -P_ID, -PIR_ID, -PIR_gene, -contains("_ES_"), - contains("_hmC_"), -contains("mC"),
         -contains("Input"), -contains("CTCF"), -contains("H3K27me3"), -contains("H3K4me1"), -contains("RNA"), -contains("P_"),
         -contains("H3K4me3"), -contains("CHiC"), -cluster_group, P_gene) %>% 
  mutate(ATAC_diff_DNMT = PIR_ATAC_DNMT_KO_EpiLC_D2 - PIR_ATAC_DNMT_WT_EpiLC_D2,
         K27ac_diff_DNMT = PIR_H3K27ac_DNMT_KO_EpiLC_D2 - PIR_H3K27ac_DNMT_WT_EpiLC_D2,
         ATAC_diff_TET = PIR_ATAC_TET_KO_EpiLC_D2 - PIR_ATAC_TET_WT_EpiLC_D2,
         K27ac_diff_TET = PIR_H3K27ac_TET_KO_EpiLC_D2 - PIR_H3K27ac_TET_WT_EpiLC_D2) %>% 
  select(interaction_ID, EM_cluster_group, Louvain_cluster_group, P_gene, contains("diff")) -> mark_differences

head(mark_differences)
```

Defining genes of interest

```{r}
genes_of_interest <- c("Tal1", "Nfe2", "Runx1", "Klf1", "Lyl1", "Gata1")
```


### EM clustering

```{r}
mark_differences %>% 
  select(-Louvain_cluster_group) %>% 
  filter(P_gene %in% genes_of_interest) -> blood_PIR_differences_EM

head(blood_PIR_differences_EM)
```

Supp Figure 8d EM (ATAC, TET)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  ggplot(aes(EM_cluster_group, ATAC_diff_TET, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("log2 fold difference ATAC\n (TET TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> SuppFig8d_EM

SuppFig8d_EM
```

Supplemental Figure 8e EM (H3K27ac, TET)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  ggplot(aes(EM_cluster_group, K27ac_diff_TET, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("log2 fold difference H3K27ac\n (TET TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> SuppFig8e_EM

SuppFig8e_EM
```
Supplemental Figure 8f EM (ATAC, DNMT)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  ggplot(aes(EM_cluster_group, ATAC_diff_DNMT, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("log2 fold difference ATAC\n (DNMT TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> SuppFig8f_EM

SuppFig8f_EM
```

Supplemental Figure 8g EM (H3K27ac, DNMT)

```{r, fig.width= 5, fig.height=5}
blood_PIR_differences_EM %>% 
  #filter (H3K27ac_diff_DNMT > -10) %>% 
  ggplot(aes(EM_cluster_group, K27ac_diff_DNMT, colour = P_gene)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_jitter(height = 0, width = 0.1) +
  scale_colour_manual(values = blood_genes_colours) +
  ylab("log2 fold difference H3K27ac\n (DNMT TKO - WT)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 12), legend.title = element_blank()) -> SuppFig8g_EM

SuppFig8g_EM
```
Exporting

```{r}
ggsave("../output/plots/SuppFigure8d_ATAC_blood_enhancers_EM.png", SuppFig8d_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/SuppFigure8d_ATAC_blood_enhancers_EM.svg", SuppFig8d_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/SuppFigure8f_ATAC_blood_enhancers_EM.png", SuppFig8f_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/SuppFigure8f_ATAC_blood_enhancers_EM.svg", SuppFig8f_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/SuppFigure8e_K27ac_blood_enhancers_EM.png", SuppFig8e_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/SuppFigure8e_K27ac_blood_enhancers_EM.svg", SuppFig8e_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/SuppFigure8g_K27ac_blood_enhancers_EM.png", SuppFig8g_EM, width = 5, height = 5, units = "in")
ggsave("../output/plots/SuppFigure8g_K27ac_blood_enhancers_EM.svg", SuppFig8g_EM, width = 5, height = 5, units = "in")
```



