---
title: "Figure 3: enhancer marks in KOs"
output: html_notebook
---

This script will create Supplemental Figure 5e of the manuscript using the EM and Louvain clusters which were identified during the paper revision process.

Defining source

```{r}
source("cluster_analysis_functions.R")
```
Importing data

```{r}
read_tsv("../data/interaction_pairs_uncapped.txt") -> magic_table_uncapped
```


## CTCF 

### EM clusters

#### cluster 10 and 19

This is the so far the only figure in which we use individual clusters instead of cluster groups. Reason is that cluster 10 was grouped with other active P - lowmarked PIR clusters so CTCF changes would be diluted.


Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_ID == 10 | EM_cluster_ID == 19) %>% 
  select(interaction_ID, EM_cluster_ID, contains("CTCF")) %>% 
  select(interaction_ID, EM_cluster_ID, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) -> CTCF_EM

head(CTCF_EM)
```

Plotting CTCF for cluster 10 and 19 - EM clusters

```{r, fig.width=2.8, fig.height=4}
CTCF_EM %>% 
  mutate(EM_cluster_ID = case_when(EM_cluster_ID == 10 ~ "cluster 10",
                                      EM_cluster_ID == 19 ~ "cluster 19")) %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("EM clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ EM_cluster_ID) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> CTCF_violins_EM

CTCF_violins_EM
```
 
### Louvain clusters

#### cluster 1 and 19, cluster 15

This is the so far the only figure in which we use individual clusters instead of cluster groups. Reason is that cluster 15 was grouped with other active P - lowmarked PIR clusters so CTCF changes would be diluted.


Filtering the magic table for these categories

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_ID == 1 | Louvain_cluster_ID == 19 | Louvain_cluster_ID == 15) %>% 
  select(interaction_ID, Louvain_cluster_ID, contains("CTCF")) %>% 
  select(interaction_ID, Louvain_cluster_ID, contains("EpiLC_D2")) %>% 
  pivot_longer(3:last_col(), names_to = "sample", values_to = "log2RPKM") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-mark, -stage, -day) %>% 
  mutate(Louvain_cluster_ID = case_when(Louvain_cluster_ID == 1 ~ "cluster 1/19",
                                        Louvain_cluster_ID == 19 ~ "cluster 1/19",
                                        Louvain_cluster_ID == 15 ~ "cluster 15")) %>% 
  mutate(Louvain_cluster_ID = factor(Louvain_cluster_ID, levels = c("cluster 15", "cluster 1/19"))) -> CTCF_Louvain

head(CTCF_Louvain)
```

Plotting CTCF for cluster 1 and 19, cluster 15 - Louvain clusters

```{r, fig.width=2.8, fig.height=4}
CTCF_Louvain %>% 
  mutate(line = case_when(line == "DNMT" ~ "Dnmt",
                          line == "TET" ~ "Tet")) %>% 
  mutate(genotype = factor(genotype, levels = c("WT", "KO"))) %>% 
  unite(line, genotype, col = "condition", sep = " ") %>% 
  mutate(condition = factor(condition, levels = c("Dnmt WT", "Dnmt KO", "", "Tet WT", "Tet KO"))) %>% 
  ggplot(aes(condition, log2RPKM, fill = condition)) +
  ylim(0,10) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ggtitle("Louvain clustering") +
  scale_fill_manual(values = c(DNMT_Epi_WT_colour, DNMT_Epi_KO_colour, TET_Epi_WT_colour, TET_Epi_KO_colour)) +
  scale_x_discrete(drop=FALSE) +
  facet_grid(fragment ~ Louvain_cluster_ID) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.background = element_rect(fill = "white"), axis.ticks.x = element_blank(), strip.text = element_text(size = 9),
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 10), plot.title = element_text(size = 12, hjust = 0.5)) -> CTCF_violins_Louvain

CTCF_violins_Louvain
```

 
Exporting

```{r}
ggsave("../output/plots/SuppFigure5e_EM.png", CTCF_violins_EM, width = 2.8, height = 4, units = "in")
ggsave("../output/plots/SuppFigure5e_EM.svg", CTCF_violins_EM, width = 2.8, height = 4, units = "in")
ggsave("../output/plots/SuppFigure5e_Louvain.png", CTCF_violins_Louvain, width = 2.8, height = 4, units = "in")
ggsave("../output/plots/SuppFigure5e_Louvain.svg", CTCF_violins_Louvain, width = 2.8, height = 4, units = "in")
```

