---
title: "Suppl Figure 4 - ES to Epi transition"
output:
  html_document:
    df_print: paged
---

This script will create the panels of Supplemental Figure 4c - h of the manuscript using the EM and Louvain clusters which were identified during the paper revision process.

Defining source

```{r}
source("cluster_analysis_functions.R")
```
Importing data

```{r}
read_tsv("../data/interaction_pairs_uncapped.txt") -> magic_table_uncapped
```

# ES to EpiLC transition (TET) - EM - clustering


as per the new clustering we are now looking for cluster groups
polycomb P - unmarked PIR
bivalent P - poised E
active P - active E
inactive P - primed E

```{r}
magic_table_uncapped %>% 
  filter(EM_cluster_group == "polycomb P - unmarked PIR" | EM_cluster_group == "bivalent P - poised E" |
           EM_cluster_group == "active P - active E" | EM_cluster_group == "inactive P - primed E") %>% 
  select(interaction_ID, EM_cluster_group, contains("WT"), -contains("DNMT"), -contains("Input"), -contains("ATAC"), -contains("CTCF"), 
         -contains("mC"), -contains("RNA"), -contains("H3K4me3")) %>% 
  rename_with(~gsub("CHiC", "P_CHiC", .)) %>% 
  pivot_longer(3:last_col(), names_to = "sample") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-line, -genotype, -day) -> SuppFigure4_EM

head(SuppFigure4_EM)
```

## Supp Figure 4c (H3K27me3) - EM-clustering

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_EM %>% 
  filter(mark == "H3K27me3") %>% 
  filter(EM_cluster_group == "polycomb P - unmarked PIR" | EM_cluster_group == "bivalent P - poised E") %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "polycomb P - unmarked PIR" ~ "Repressed",
                                       EM_cluster_group == "bivalent P - poised E" ~ "Poised")) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Repressed", "Poised"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K27me3 (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4c_EM

SuppFigure4c_EM
```


## Supp Figure 4d (H3K4me1)

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_EM %>% 
  filter(mark == "H3K4me1") %>% 
  filter(EM_cluster_group == "polycomb P - unmarked PIR" | EM_cluster_group == "bivalent P - poised E") %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "polycomb P - unmarked PIR" ~ "Repressed",
                                       EM_cluster_group == "bivalent P - poised E" ~ "Poised")) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Repressed", "Poised"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K4me1 (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4d_EM

SuppFigure4d_EM
```
## supp figure 4e 

```{r, fig.height=2.5, fig.width=2.1}
SuppFigure4_EM %>% 
  filter(mark == "CHiC") %>% 
  filter(EM_cluster_group == "polycomb P - unmarked PIR" | EM_cluster_group == "bivalent P - poised E") %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "polycomb P - unmarked PIR" ~ "Repressed",
                                       EM_cluster_group == "bivalent P - poised E" ~ "Poised")) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Repressed", "Poised"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("HiC (CHiCAGO)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4e_EM

SuppFigure4e_EM
```
## Supp Figure 4f (H3K4me1)

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_EM %>% 
  filter(mark == "H3K4me1") %>% 
  filter(EM_cluster_group == "active P - active E" | EM_cluster_group == "inactive P - primed E") %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "active P - active E" ~ "Active",
                                       EM_cluster_group == "inactive P - primed E" ~ "Primed")) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Primed", "Active"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K4me1 (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4f_EM

SuppFigure4f_EM
```

## Supp Figure 4g (H3K27ac)

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_EM %>% 
  filter(mark == "H3K27ac") %>% 
  filter(EM_cluster_group == "active P - active E" | EM_cluster_group == "inactive P - primed E") %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "active P - active E" ~ "Active",
                                       EM_cluster_group == "inactive P - primed E" ~ "Primed")) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Primed", "Active"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K27ac (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4g_EM

SuppFigure4g_EM
```

## supp figure 4h 

```{r, fig.height=2.5, fig.width=2.1}
SuppFigure4_EM %>% 
  filter(mark == "CHiC") %>% 
  filter(EM_cluster_group == "active P - active E" | EM_cluster_group == "inactive P - primed E") %>% 
  mutate(EM_cluster_group = case_when(EM_cluster_group == "active P - active E" ~ "Active",
                                       EM_cluster_group == "inactive P - primed E" ~ "Primed")) %>% 
  mutate(EM_cluster_group = factor(EM_cluster_group, levels = c("Primed", "Active"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("HiC (CHiCAGO)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ EM_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4h_EM

SuppFigure4h_EM
```
 
## Exporting the EM supp fig 4 plots

```{r}
ggsave("../output/plots/SuppFigure4c_EM.png", SuppFigure4c_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4c_EM.svg", SuppFigure4c_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4d_EM.png", SuppFigure4d_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4d_EM.svg", SuppFigure4d_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4f_EM.png", SuppFigure4f_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4f_EM.svg", SuppFigure4f_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4g_EM.png", SuppFigure4g_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4g_EM.svg", SuppFigure4g_EM, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4e_EM.png", SuppFigure4e_EM, width = 2.1, height = 2.5, units = "in")
ggsave("../output/plots/SuppFigure4e_EM.svg", SuppFigure4e_EM, width = 2.1, height = 2.5, units = "in")
ggsave("../output/plots/SuppFigure4h_EM.png", SuppFigure4h_EM, width = 2.1, height = 2.5, units = "in")
ggsave("../output/plots/SuppFigure4h_EM.svg", SuppFigure4h_EM, width = 2.1, height = 2.5, units = "in")
```

 

# ES to EpiLC transition (TET) - Louvain - clustering


as per the new clustering we are now looking for cluster groups
polycomb P - unmarked PIR
bivalent P - poised E
active P - active E
inactive P - primed E

```{r}
magic_table_uncapped %>% 
  filter(Louvain_cluster_group == "polycomb P - unmarked PIR" | Louvain_cluster_group == "bivalent P - poised E" |
           Louvain_cluster_group == "active P - active E" | Louvain_cluster_group == "inactive P - primed E") %>% 
  select(interaction_ID, Louvain_cluster_group, contains("WT"), -contains("DNMT"), -contains("Input"), -contains("ATAC"), -contains("CTCF"), 
         -contains("mC"), -contains("RNA"), -contains("H3K4me3")) %>% 
  rename_with(~gsub("CHiC", "P_CHiC", .)) %>% 
  pivot_longer(3:last_col(), names_to = "sample") %>% 
  separate(sample, into = c("fragment", "mark", "line", "genotype", "stage", "day")) %>% 
  select(-line, -genotype, -day) -> SuppFigure4_Louvain

head(SuppFigure4_Louvain)
```

## Supp Figure 4c (H3K27me3) - Louvain-clustering

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_Louvain %>% 
  filter(mark == "H3K27me3") %>% 
  filter(Louvain_cluster_group == "polycomb P - unmarked PIR" | Louvain_cluster_group == "bivalent P - poised E") %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "polycomb P - unmarked PIR" ~ "Repressed",
                                       Louvain_cluster_group == "bivalent P - poised E" ~ "Poised")) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Repressed", "Poised"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K27me3 (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4c_Louvain

SuppFigure4c_Louvain
```


## Supp Figure 4d (H3K4me1)

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_Louvain %>% 
  filter(mark == "H3K4me1") %>% 
  filter(Louvain_cluster_group == "polycomb P - unmarked PIR" | Louvain_cluster_group == "bivalent P - poised E") %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "polycomb P - unmarked PIR" ~ "Repressed",
                                       Louvain_cluster_group == "bivalent P - poised E" ~ "Poised")) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Repressed", "Poised"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K4me1 (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4d_Louvain

SuppFigure4d_Louvain
```
## supp figure 4e 

```{r, fig.height=2.5, fig.width=2.1}
SuppFigure4_Louvain %>% 
  filter(mark == "CHiC") %>% 
  filter(Louvain_cluster_group == "polycomb P - unmarked PIR" | Louvain_cluster_group == "bivalent P - poised E") %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "polycomb P - unmarked PIR" ~ "Repressed",
                                       Louvain_cluster_group == "bivalent P - poised E" ~ "Poised")) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Repressed", "Poised"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("HiC (CHiCAGO)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4e_Louvain

SuppFigure4e_Louvain
```
## Supp Figure 4f (H3K4me1)

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_Louvain %>% 
  filter(mark == "H3K4me1") %>% 
  filter(Louvain_cluster_group == "active P - active E" | Louvain_cluster_group == "inactive P - primed E") %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "active P - active E" ~ "Active",
                                       Louvain_cluster_group == "inactive P - primed E" ~ "Primed")) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Primed", "Active"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K4me1 (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4f_Louvain

SuppFigure4f_Louvain
```

## Supp Figure 4g (H3K27ac)

```{r, fig.height=4, fig.width=2.1}
SuppFigure4_Louvain %>% 
  filter(mark == "H3K27ac") %>% 
  filter(Louvain_cluster_group == "active P - active E" | Louvain_cluster_group == "inactive P - primed E") %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "active P - active E" ~ "Active",
                                       Louvain_cluster_group == "inactive P - primed E" ~ "Primed")) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Primed", "Active"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("H3K27ac (RPKM)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4g_Louvain

SuppFigure4g_Louvain
```

## supp figure 4h 

```{r, fig.height=2.5, fig.width=2.1}
SuppFigure4_Louvain %>% 
  filter(mark == "CHiC") %>% 
  filter(Louvain_cluster_group == "active P - active E" | Louvain_cluster_group == "inactive P - primed E") %>% 
  mutate(Louvain_cluster_group = case_when(Louvain_cluster_group == "active P - active E" ~ "Active",
                                       Louvain_cluster_group == "inactive P - primed E" ~ "Primed")) %>% 
  mutate(Louvain_cluster_group = factor(Louvain_cluster_group, levels = c("Primed", "Active"))) %>% 
  mutate(stage = gsub("ES", "ESC", stage)) %>% 
  mutate(stage = factor(stage, levels = c("ESC", "EpiLC"))) %>% 
  ggplot(aes(stage, value, fill = stage)) +
  geom_violin(draw_quantiles = 0.5, scale = "width") +
  ylim(0,10) +
  ggtitle("HiC (CHiCAGO)") +
  scale_fill_manual(values = c(ES_colour, Epi_colour)) +
  facet_grid(fragment ~ Louvain_cluster_group) +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_text(size = 8, margin = margin(r=0)), 
        panel.grid = element_blank(), strip.text.y = element_blank(), strip.background = element_rect(fill = "white", colour = NA), 
        axis.ticks.x = element_blank(), strip.text = element_text(size = 8), axis.title.y = element_blank(),
        axis.text.y = element_text(size = 8), plot.title = element_text(size = 8, hjust = 0.5, vjust = -4)) -> SuppFigure4h_Louvain

SuppFigure4h_Louvain
```

## Exporting the Louvain supp fig 4 plots

```{r}
ggsave("../output/plots/SuppFigure4c_Louvain.png", SuppFigure4c_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4c_Louvain.svg", SuppFigure4c_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4d_Louvain.png", SuppFigure4d_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4d_Louvain.svg", SuppFigure4d_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4f_Louvain.png", SuppFigure4f_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4f_Louvain.svg", SuppFigure4f_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4g_Louvain.png", SuppFigure4g_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4g_Louvain.svg", SuppFigure4g_Louvain, width = 2.1, height = 4, units = "in")
ggsave("../output/plots/SuppFigure4e_Louvain.png", SuppFigure4e_Louvain, width = 2.1, height = 2.5, units = "in")
ggsave("../output/plots/SuppFigure4e_Louvain.svg", SuppFigure4e_Louvain, width = 2.1, height = 2.5, units = "in")
ggsave("../output/plots/SuppFigure4h_Louvain.png", SuppFigure4h_Louvain, width = 2.1, height = 2.5, units = "in")
ggsave("../output/plots/SuppFigure4h_Louvain.svg", SuppFigure4h_Louvain, width = 2.1, height = 2.5, units = "in")
```





